---
title: "Electric power sector transition scenarios"
author: '@ideea'
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(IDEEA)
source("R/setup.R")
```

# Configuration
```{r, eval = T}
# number of regions
nreg <- 5 
# nreg <- 32 
```

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(sf)
library(glue)
library(IDEEA)

regN <- glue("reg", nreg)
regN_off <- paste0(regN, "_off")

```

## Regions

```{r regions}
ideea_sf <- get_ideea_map(nreg, offshore = T, reg_off = T)
# quich plot
plot(ideea_sf[regN], key.width = lcm(4))

# alternative: ggplot2 version
if (F) {
  ggplot(ideea_sf) +
    geom_sf(aes(fill = reg5_off)) +
    theme_ideea_map()
}
```

## Electricity model
Building the electricity model from pre-build IDEEA modules, saved in the `IDEEA::ideea_modules` list.

```{r}
names(ideea_modules) # shows names of modules

# Import all electricity model objects for selected number of regions
names(ideea_modules$electricity) # all available models
elc <- ideea_modules$electricity[[regN]]
class(elc)
names(elc)
summary(elc)

# # Update horizon up to 2075
# horizon_2020_2070_by_10 <-
#   newHorizon(
#   name = "Y2020_2070_by_10",
#   desc = "2020 to 2070 by 10 years",
#   period = 2020:2075,
#   intervals = c(1, 5, rep(10, 15)),
#   mid_is_end = F
# )
# horizon_2020_2070_by_10
# # (optional) Change milestone year 2023 to 2025
# horizon_2020_2070_by_10@intervals # see 2nd row
# horizon_2020_2070_by_10@intervals$mid[2] <- 2025
# horizon_2020_2070_by_10@intervals # check

# horizon_2020_2070_by_20 <-
#   newHorizon(
#   name = "Y2020_2070_by_20",
#   desc = "2020 to 2070 by 20 years",
#   period = 2020:2070,
#   intervals = c(1, 5, 10, rep(20, 15)),
#   mid_is_end = T
# )
# horizon_2020_2070_by_20@intervals
# $mid[3] <- 2050

horizon_2020_2070_by_15 <- newHorizon(
  name = "Y2020_2070_by_15",
  desc = "2020 to 2070 by 20 years",
  period = 2020:2070,
  intervals = c(1, 5, 10, rep(15, 5)),
  mid_is_end = T
)

# Updating constraints (to 2070+)
NO_NEW_HYDRO <- elc$NO_NEW_HYDRO
NO_NEW_HYDRO@rhs <- data.frame(
  year = 2020:2100, # extend the constraint
  rhs = 1e-20 # close to zero (non-zero to avoid dropping)
  ) 
NO_NEW_HYDRO@for.each <- data.frame(year = 2020:2100)

NO_NEW_NUCLEAR <- elc$NO_NEW_NUCLEAR
NO_NEW_NUCLEAR@rhs <- data.frame(
  year = 2020:2100, # all range of years
  rhs = 1e-20
  ) 
NO_NEW_NUCLEAR@for.each <- data.frame(year = 2020:2100)

# A model with the default configuration
mod <- newModel(
  name = glue("IDEEA_ELC_{regN}"),
  desc = "IDEEA electricity example model",
  region = unique(ideea_sf[[regN]]),
  discount = 0.05,
  calendar = ideea_modules$calendars$calendar_d365_h24,
  horizon = ideea_modules$horizons$horizon_2020_2060_by_10,
  data = newRepository(
    name = glue("repo_electricity_{regN}"),
    elc$repo_comm, # declaration of commodities used in the model
    # supply & import
    elc$repo_supply, # primary energy supply (markets)
    elc$repo_geoccs, # CCS storage potential by region
    # Generating technologies
    elc$ECOA, # coal-fired generators
    elc$EGAS, # gas-fired generators
    elc$ENUC, # nuclear generators
    elc$EHYD, # hydro generators
    elc$ESOL_w_const, # solar generators with constraints on capacity
    # elc$ESOL_w_land, # solar generators with constraints on land
    elc$EWIN_w_const, # wind generators with constraints on capacity
    # elc$EWIN_w_land, # wind generators with constraints on land
    elc$EWIF_w_const, # wind generators with constraints on capacity
    # elc$EWIF_w_land, # wind generators with constraints on used surface area
    elc$EBIO, # biomass generators
    elc$repo_ccstechs, # CCS technologies
    # capacity (weather) factors
    elc$WSOL, # solar capacity factor
    elc$WWIN, # onshore wind capacity 
    elc$WWIF, # offshore wind capacity 
    elc$WHYD, # hydro capacity factor
    # battery
    elc$STG_BTR, # battery storage with yearly cycle
    # STG_BTR_daily = STG_BTR_daily, # battery storage with daily cycle
    # transmission
    elc$repo_transmission_ac, # HVAC transmission network
    elc$repo_transmission_dc, # HVDC transmission network with converters
    # assumptions / policies
    NO_NEW_HYDRO, # no new hydro capacity
    NO_NEW_NUCLEAR, # no new nuclear capacity
    # elc$NO_NEW_WIND, # no new wind capacity
    # elc$NO_NEW_SOLAR, # no new solar capacity
    # demand
    elc$DEMELC_BY, # electricity demand in the base year
    elc$DEMELC_2X # electricity demand growth
  )
)
## Updating carbon policy to net zero by 2070
CO2_CAP_2070 <- ideea_modules$electricity[[regN]]$CO2_CAP
CO2_CAP_2070@rhs
CO2_CAP_2070@rhs$year[CO2_CAP_2070@rhs$year == 2060] <- 2070
CO2_CAP_2070@rhs

```

## Input check

### Wind clusters
<!-- ToDo: done -->
<!-- - add cluster information (N and tol) to the repository to avoid ambiguity   -->
<!-- - write a function for quick plotting of clustered data (wind/solar) - DONE -->

```{r}
names(elc$EWIN_w_const@data)
message("Number of regions: ", nreg)
```

### Plot wind capacity factor by cluster
<!-- ToDO: done -->
<!-- - debug gif function -->
<!-- - share IDEEA extra-repository -->
```{r}
# resource <- "win"; 
win_tol <- 0.05
calendar_1day_per_month <- 
  ideea_modules$calendars$calendar_d365_h24_subset_1day_per_month
  
slices_1day_per_month <- calendar_1day_per_month@timetable$slice

# ideea_sf <- get_ideea_map(nreg = nreg, offshore = T, islands = T)
ideea_win_cl_sf <- get_ideea_cl_sf(resource = "win", tol = win_tol)
plot(ideea_win_cl_sf["wcf_100m"])

win_cf_data <- get_ideea_cf("win", tol = win_tol, year = 2019)
win_cf_data
cf_name <- "wcf_100m"

# plot
wind_snapshot <- ideea_snapshot_cf(win_cf_data, ideea_win_cl_sf, ideea_sf, 
                  cf_name = cf_name, 
                  slice = slices_1day_per_month[100])
wind_snapshot
ggsave("images/wind_snapshot.png", wind_snapshot, scale = 1.5,
       width = 5, height = 5)

# gif
# ?ideea_gif_cf
ideea_gif_cf(x = win_cf_data, 
             ideea_cl_sf = ideea_win_cl_sf,
             ideea_sf = ideea_sf,
             cf_name = cf_name,
             slice = slices_1day_per_month[1:24],
             fps = 5, gif.width = 864, gif.height = 864,
             # fps = 5, gif.width = 432, gif.height = 432,
             filename = glue("wind_capacity_factors_sample.gif"))
```

### Plot solar capacity factors by cluster
```{r}
sol_tol <- 0.01
ideea_sol_cl_sf <- get_ideea_cl_sf(resource = "sol", tol = sol_tol)
sol_cf_data <- get_ideea_cf("sol", tol = sol_tol, year = 2019)
sol_cf_data
# plot
solar_snapshot <- ideea_snapshot_cf(sol_cf_data, ideea_sol_cl_sf, ideea_sf, 
                  cf_name = "scf_tl")
solar_snapshot
ggsave("images/solar_snapshot.png", solar_snapshot, scale = 1.5,
       width = 5, height = 5)
library(patchwork)
sol_win <- solar_snapshot + wind_snapshot
ggsave("images/solar_wind_snapshot.png", sol_win, scale = 1.5,
       width = 8.5, height = 5)

# gif
# sol_cf_data_EAST <- filter(sol_cf_data, reg5 == "EAST")
ideea_gif_cf(
  x = sol_cf_data,
  ideea_cl_sf = ideea_sol_cl_sf,
  ideea_sf = ideea_sf,
  cf_name = "scf_tl",
  slice = slices_1day_per_month[1:24],
  fps = 5, gif.width = 864, gif.height = 864,
  # fps = 5, gif.width = 332, gif.height = 432,
  filename = glue("sol_capacity_factors_sample_EAST.gif"))

```


### (optional) cluster locations
```{r, eval=FALSE}
fig.wind <- ggplot() +
  geom_sf(data = ideea_sf$geometry) +
  geom_sf(data = ideea_win_cl_sf, aes(fill = wcf_100m), color = "grey15") +
  scale_fill_viridis_c(option = "D", name = "CF") +
  labs(title = "Wind, 100m hub") +
  theme_ideea_map() +
  theme(plot.title = element_text(hjust = 0.5))

fig.sol <- ggplot() +
  geom_sf(data = ideea_sf$geometry) +
  geom_sf(data = ideea_sol_cl_sf, aes(fill = scf_tl), color = "grey85") +
  scale_fill_viridis_c(option = "A", name = "CF") +
  labs(title = "Solar, 1-axis tracking") +
  theme_ideea_map() +
  theme(plot.title = element_text(hjust = 0.5))

library(patchwork)
sol_win <- fig.sol + fig.wind
ggsave("images/solar_wind_cf.png", sol_win, scale = 1.5,
       width = 8.5, height = 5)

```


```{r}

EXPELC <- newExport(
  name = "EXPELC",
  desc = "Pseudo-export to capture curtailed generation",
  commodity = "ELC",
  unit = "cr.₹/GWh",
  exp = list(
    price = 0.001 # "cr.INR/GWh" 
    # convert("cr.₹/GWh", "cents/kWh", .001) # Check
    # convert("cr.₹/GWh", "₹/kWh", .001) # Check
  )
)

IMPELC <- newImport(
  name = "IMPELC",
  desc = "Unserved load penalty",
  commodity = "ELC",
  unit = "cr.₹/GWh",
  imp = list(
    price = 1e4 # high marginal price (penalty) for unserved load
    # convert("USD/kWh", "MUSD/GWh", 1e4)
  )
)
```



## Scenarios
Let's consider two scenarios for the electricity sector in 2050. The first scenario is a reference scenario (**REF**) without any policy intervention. The second scenario is a scenario with a carbon emissions constraint (**CAP**), targeting zero emissions by 2060 (linear reduction from 2030).  

Here we solve both scenarios for 2050 with full 8760 hours resolution and a subset of 1 day per month, 288 hours (24 * 12) in total. The subset scenario is relatively fast to solve and can be used for drafting and testing purposes. Though it does not provide a full picture of intermittent renewable generation and can downplay the role of storage and transmission.


### Reference scenario
The **REF** scenario is a business-as-usual that includes all the existing technologies that are already in the model. 

#### Solving for 288 hours
The 288 hours scenario can be solved with open source solvers like HiGHS or Cbc in minutes. Here we use the HiGHS solver with the barrier method and Julia interface.
```{r, eval=FALSE}
# (optional) set default solver
# set_default_solver(solver_options$julia_highs_parallel)
# set_progress_bar(type = "progress") # output progress bar in the console

# Sub-setting and interpolating of the model parameters
scen_REF_sub <- interpolate_model(
  mod,
  name = "REF3_sub", # indicating 'subset'
  calendar = ideea_modules$calendars$calendar_d365_h24_subset_1day_per_month,
  horizon = ideea_modules$horizons$horizon_2070
)
summary(scen_REF_sub)

# Writing the model script and data
scen_REF_sub <- write_sc(
  scen_REF_sub
  # solver = solver_options$julia_cplex
  # solver = solver_options$julia_highs_barrier
  # solver = solver_options$julia_highs
)
scen_REF_sub@status # info
scen_REF_sub@misc$tmp.dir # temporary directory

# Solving the scenario
scen_REF_sub <- solve_scenario(scen_REF_sub, wait = T)

# Reading the results
scen_REF_sub <- read(scen_REF_sub)
summary(scen_REF_sub)

# (optional) saving the scenario on disk
scen_REF_sub <- save_scenario(scen_REF_sub)
size(scen_REF_sub)

```

### Carbon emissions constraint scenario
The **CAP** scenario is a scenario with a carbon emissions constraint. Here we use pre-defined CO2 emissions constraint on the national level, assuming emissions trading. The constraint is saved in `ideea_modules[[regN]]$CO2_CAP` object.

#### Solving for 288 hours
```{r, eval=FALSE}
# Policy constraint 
# CO2_CAP <- ideea_modules$electricity[[regN]]$CO2_CAP

# Sub-setting and interpolating of the model parameters
scen_CAP_sub <- interpolate(
  mod,
  name = "CAP3_sub", # indicating 'subset'
  CO2_CAP_2070, # carbon emissions constraint
  calendar = ideea_modules$calendars$calendar_d365_h24_subset_1day_per_month,
  horizon = ideea_modules$horizons$horizon_2070
)

# Writing the model script and data
scen_CAP_sub <- write_sc(
  scen_CAP_sub
  # solver = solver_options$gams_gdx_cplex_parallel
  # solver = solver_options$julia_highs_barrier
  # solver = julia_highs_crossover_on
)

# Solving the scenario
scen_CAP_sub <- solve_scenario(scen_CAP_sub, wait = T)

# Reading the results
scen_CAP_sub <- read(scen_CAP_sub)
summary(scen_CAP_sub)

# (optional) saving the scenario on disk
scen_CAP_sub <- save_scenario(scen_CAP_sub)

```


## Solving for 8760 hours
The full-year scenario requires more computational resources and time. The generation of the problem also varies by the software. Here we use Python/Pyomo and CPLEX solver. The one-year scenario can be solved within an hour on a consumer-grade desktop or laptop. Barrier algorithm benefits from parallel processing and is normally faster on multi-core CPUs.  

### REF
```{r}
# Interpolating the model parameters
scen_REF <- interpolate(
  mod,
  name = "REF3",
  horizon = horizon_2020_2070_by_15
)

# Writing the model script and data
scen_REF <- write_sc(scen_REF)

# Solving the scenario
scen_REF <- solve_scenario(scen_REF, wait = F)
# save(scen_REF, file = file.path(ideea_scenarios(), "scen_REF.RData"))
# rm(scen_REF); gc()
(load(file.path(ideea_scenarios(), "scen_REF.RData")))

# Reading the results
scen_REF <- read(scen_REF)
summary(scen_REF)

# (optional) saving the scenario on disk
scen_REF <- save_scenario(scen_REF)

```

### CAP
Similarly, the full-year scenario can be solved with the same solver and settings as **REF**.
```{r}
# Interpolating the model parameters
scen_CAP <- interpolate_model(
  mod, IMPELC, EXPELC,
  name = "CAP3",
  CO2_CAP_2070, # carbon emissions constraint
  horizon = horizon_2020_2070_by_15
)

# Writing the model script and data
scen_CAP <- write_script(scen_CAP)
  # replace default solver
  # solver_options$pyomo_cplex_barrier
  # solver_options$julia_cplex_barrier
  # solver = solver_options$gams_gdx_cplex_barrier
  # solver = solver_options$gams_gdx_cplex_parallel
  # solver = solver_options$julia_highs_barrier
  # solver = solver_options$julia_highs_parallel
# )

# Solving the scenario
# save(scen_CAP, file = file.path(ideea_scenarios(), "scen_CAP.RData"))
scen_CAP <- solve_scenario(scen_CAP, wait = F)
# installr::os.sleep()

# Reading the results
scen_CAP <- read(scen_CAP)
summary(scen_CAP)
isInMemory(scen_CAP) 

# (optional) saving the scenario on disk
scen_CAP <- save_scenario(scen_CAP)
isInMemory(scen_CAP) # FALSE means the scenario is saved on disk
summary(scen_CAP) # check RAM size 
# installr::os.sleep()

```


## Results
... an example of processing some results ...

```{r}
all_slices <- scen_CAP@settings@calendar@timetable$slice

# (very) quick plots to use
?ideea_snapshot
ideea_snapshot(scen_REF, YEAR = 2050, SLICE = all_slices[1:24])
ideea_snapshot(scen_CAP, YEAR = 2050, SLICE = all_slices[1:24])
scen_cap_operation_data <- ideea_snapshot(scen_CAP, YEAR = 2050, 
               SLICE = all_slices, 
               return_data = T)
fwrite(scen_cap_operation_data, file = "tmp/scen_cap_operation_data.csv")

# ToDo: 
# - add functionality to plot several scenarios
# - fix colors / palette for fuel mix and techs

```



```{r}
# Combine scenarios in a named or unnamed list
# (in the case of unnamed list, scenario names are used)
sns <- list(scen_REF_sub, scen_CAP_sub)
# sns <- list(scen_REF_sub, scen_REF, scen_CAP_sub, scen_CAP)
# sns <- scen_REF_sub

```

Use `getData()` function to retrieve variables and parameters from solved scenario objects. See `?getData` for more details.
```{r}
# Objective function value (total discounted cost)
getData(sns, "vObjective", merge = T)
```
The "subset" time-slices scenarios normally show lower objective. 

Compare technological capacity by scenario.
```{r}
vTechCap <- getData(sns, "vTechCap", merge = T, process = T, 
                    tech_ = "^E",
                    digits = 2, drop.zeros = T) |> 
  drop_process_cluster() |> # removes cluster indicators from process names
  drop_process_vintage() |> # removes vintage indicators from process names
  group_by(scenario, process, year) |> # group and aggregate for figures/tables
  summarise(GW = sum(value), .groups = "drop") |>
  filter(!grepl("EDC", process)) |>
  as.data.table()
vTechCap |> pivot_wider(names_from = process, values_from = GW)

ggplot(vTechCap) +
  geom_bar(aes(x = scenario, y = GW, fill = process), stat = "identity") +
  scale_fill_viridis_d(option = "H", direction = -1) +
  theme_bw() 

vEmsFuelTot <- getData(sns, "vEmsFuelTot", merge = T) 

ggplot(vEmsFuelTot) +
  geom_bar(aes(x = scenario, y = value), stat = "identity", fill = "brown") +
  # scale_fill_viridis_d(option = "H", direction = -1) +
  theme_bw() 

```

