---
title: "Electric power sector transition scenarios (with new data)"
author: '@ideea'
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(IDEEA)
source("R/setup.R")
TZ <- "Asia/Kolkata"
```

# Configuration
```{r, eval = T}
# number of regions
nreg <- 5 
# nreg <- 32 
```

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(sf)
library(glue)
library(IDEEA)

regN <- glue("reg", nreg)
regN_off <- paste0(regN, "_off")

```

## Regions

```{r regions}
ideea_sf <- get_ideea_map(nreg, offshore = T, reg_off = T)
# quich plot
plot(ideea_sf[regN], key.width = lcm(4))

# alternative: ggplot2 version
if (F) {
  ggplot(ideea_sf) +
    geom_sf(aes(fill = reg5_off)) +
    theme_ideea_map()
}
```

## Electricity model
Building the electricity model from pre-build IDEEA modules, saved in the `IDEEA::ideea_modules` list.

```{r}
names(ideea_modules) # shows names of modules

# Import all electricity model objects for selected number of regions
names(ideea_modules$electricity) # all available models
elc <- ideea_modules$electricity[[regN]]
class(elc)
names(elc)
summary(elc)

```

## Using new data 
(from Oct-27-2025)

### Load curve
```{r}
new_loadcurve <- fread("data/load_curve_2025_r7.csv.gz")
new_loadcurve$datetime[1] # check timezone - csv files not always store TZ
# update tz
new_loadcurve$datetime <- with_tz(new_loadcurve$datetime, tzone = TZ) 
new_loadcurve$DATE <- as.Date(new_loadcurve$datetime, tz = TZ)
# Add time-slices
new_loadcurve$timeslice <- dtm2tsl(new_loadcurve$datetime)
new_loadcurve <- new_loadcurve |> 
  filter(year == 2024) |>
  mutate(
    timeslice = dtm2tsl(datetime, d366.as.na = TRUE), # keep 365 days
    GWh = value / 1e3 # convert to the model units
  ) |>
  filter(!is.na(timeslice))
nrow(new_loadcurve) / 5  # check 8760, e.g. all data is present

# updating the base year demand
DEMELC_BY <- elc$DEMELC_BY

DEMELC_BY@dem |> head()

DEMELC_BY_2024 <- update(
  DEMELC_BY, 
  desc = "Load curve 2024, source Niti Aayog",
  dem = data.frame(
    region = new_loadcurve$reg7, # reg7 is equal to reg5 when no offshores
    # year = 2024, #  the same for all years
    slice = new_loadcurve$timeslice,
    dem = new_loadcurve$GWh
  ))

DEMELC_BY_2024@dem |> head()
# save(new_loadcurve, file = "data/new_load_tmp.RData")

elc$DEMELC_BY@dem |> filter(slice == "d010_h01", region == "NORTH") 
elc$DEMELC_2X@dem |> filter(slice == "d010_h01", region == "NORTH") 

# assume 3x growth by 2050 from 2024 base year and 4x by 2070
DEMELC_2X_2024 <- elc$DEMELC_2X

dem_growth_2025 <- DEMELC_BY_2024@dem |> 
  mutate(
    year = 2025,
    dem = 0. # growth component is zero in base year
  )

dem_growth_2050 <- DEMELC_BY_2024@dem |>
  mutate(
    year = 2050,
    dem = 2 * dem # additional demand 2x by 2050 (total 3x)
  )

dem_growth_2070 <- DEMELC_BY_2024@dem |>
  mutate(
    year = 2070,
    dem = 3 * dem # additional demand 3x by 2070 (total 4x)
  )

DEMELC_2X_2024 <- update(
  DEMELC_2X_2024,
  desc = "Demand growth 2025, 2050, 2070 from 2024 base year",
  dem = rbindlist(list(dem_growth_2025, dem_growth_2050, dem_growth_2070))
)


```

### Pre-existing capacity
```{r}
new_capacity <- fread("data/gen_capacity_2025_r7.csv.gz") |>
  filter(year == max(year)) |>
  filter(month == max(month)) # the latest data

new_capacity$source |> unique() # available sources
```

```{r}
ggplot(new_capacity) +
  geom_col(aes(reg7, value/1e3, fill = source)) +
  scale_fill_viridis_d(option = "H", direction = -1) +
  theme_bw()
```


#### Coal plants
```{r}
ECOA_2025 <- elc$ECOA
names(ECOA_2025)

# ultra-supercritical - no capacity in 2020, ToDo: check for 2025
ECOA_2025@data$ECOAULT_2020@capacity 

# supercritical (basic tech) 
ECOA_2025@data$ECOASUP_2020@capacity

coa_2025 <- new_capacity |> filter(grepl("Coal", source, ignore.case = T))
coa_2025

coa_2050 <- coa_2025 |>
  mutate(
    year = 2050,
    value = 0.5 * value) # assume 50% of retirement by 2050

coa_2070 <- coa_2025 |>
  mutate(
    year = 2070,
    value = 0.0 * value) # assume 100% of retirement by 2070

coa_cap <- rbindlist(list(coa_2025, coa_2050, coa_2070))

# update tech
ECOA_2025@data$ECOASUP_2025 <- update(
  ECOA_2025@data$ECOASUP_2020,
  name = "ECOASUP_2025",
  desc = "Updated coal capacity, 2025",
  capacity = data.frame(
    region = coa_cap$reg7,
    year = coa_cap$year,
    stock = round(coa_cap$value / 1e3, 1) # drop crumbs
  )
)
ECOA_2025@data$ECOASUP_2025@capacity
ECOA_2025@data$ECOAULT_2020 <- NULL # drop to avoid installation
ECOA_2025@data$ECOASUP_2020 <- NULL # replaced with 2025

```

#### Nat Gas
(assuming Oil & Gas is Gas only)
```{r}
EGAS_2025 <- elc$EGAS
EGAS_2025@data$ENGCC_2020@capacity # assign all here
EGAS_2025@data$ENGOC_2020@capacity # does not exist

gas_2025 <- new_capacity |> filter(grepl("Gas", source, ignore.case = T))
gas_2025 
# the numbers in 2025 are smaller and quite close to existing - skip for now

```

#### Nuclear
```{r}
elc$ENUC@data$ENUC_2020@capacity # existing capacity
nuc_2025 <- new_capacity |> filter(grepl("Nuclear", source, ignore.case = T))
nuc_2025

nuc_2070 <- nuc_2025 |>
  mutate(
    year = 2070,
    value = 1 * value) # no retirement assumed

nuc_cap <- rbindlist(list(nuc_2025, nuc_2070))

# update tech
ENUC_2025 <- elc$ENUC
ENUC_2025@data$ENUC_2025 <- update(
  ENUC_2025@data$ENUC_2020,
  name = "ENUC_2025",
  desc = "Updated nuclear capacity, 2025",
  capacity = data.frame(
    region = nuc_cap$reg7,
    year = nuc_cap$year,
    stock = round(nuc_cap$value / 1e3, 1) # drop crumbs
  )
)
ENUC_2025@data$ENUC_2025@capacity
ENUC_2025@data$ENUC_2020 <- NULL # replaced with 2025

```

#### Hydro
```{r}
elc$EHYD |> names()
EHYD_2025 <- elc$EHYD
EHYD_2025@data$EHYD_2020@capacity # existing capacity
hyd_2025 <- new_capacity |> filter(grepl("Hydro", source, ignore.case = T)) |>
  # aggregate small and large hydro
  group_by(reg7, year) |>
  summarise(value = sum(value), .groups = "drop")

hyd_2050 <- hyd_2025 |>
  mutate(
    year = 2050,
    value = 1 * value) # no retirement assumed

hyd_2100 <- hyd_2025 |>
  mutate(
    year = 2100,
    value = 1 * value) # no retirement assumed

hyd_cap <- rbindlist(list(hyd_2025, hyd_2050, hyd_2100))

# update tech
EHYD_2025@data$EHYD_2025 <- update(
  EHYD_2025@data$EHYD_2020,
  name = "EHYD_2025",
  desc = "Updated hydro capacity, 2025",
  capacity = data.frame(
    region = hyd_cap$reg7,
    year = hyd_cap$year,
    stock = round(hyd_cap$value / 1e3, 1) # drop crumbs
  )
)

EHYD_2025@data$EHYD_2025@capacity
EHYD_2025@data$EHYD_2020 <- NULL # replaced with 2025
```

#### Biomass
```{r}
elc$EBIO |> names()
EBIO_2025 <- elc$EBIO
EBIO_2025@data$EBIO_2020@capacity # existing capacity
bio_2025 <- new_capacity |> filter(grepl("Bio", source, ignore.case = T))
bio_2025

bio_2050 <- bio_2025 |>
  mutate(
    year = 2050,
    value = 1 * value) # no retirement assumed

bio_2070 <- bio_2025 |>
  mutate(
    year = 2070,
    value = 0 * value) # full retirement assumed

bio_cap <- rbindlist(list(bio_2025, bio_2050, bio_2070))

# update tech
EBIO_2025@data$EBIO_2025 <- update(
  EBIO_2025@data$EBIO_2020,
  name = "EBIO_2025",
  desc = "Updated biomass capacity, 2025",
  capacity = data.frame(
    region = bio_cap$reg7,
    year = bio_cap$year,
    stock = round(bio_cap$value / 1e3, 1) # drop crumbs
  )
)
EBIO_2025@data$EBIO_2025@capacity
EBIO_2025@data$EBIO_2020 <- NULL # replaced with 2025

```


#### Solar
```{r}
elc$ESOL_w_const |> names()
ESOL_2025 <- elc$ESOL_w_const
ESOL_2025@data$ESOL_2020_CL01@capacity # existing capacity
ESOL_2025@data$ESOL_2020_CL02@capacity # existing capacity
ESOL_2025@data$ESOL_2020_CL03@capacity # existing capacity

sol_2025 <- new_capacity |> filter(grepl("Solar", source, ignore.case = T))
sol_2025
sol_2050 <- sol_2025 |>
  mutate(
    year = 2050,
    value = 1 * value) # no retirement assumed

sol_2070 <- sol_2025 |>
  mutate(
    year = 2070,
    value = 0 * value) # full retirement assumed

sol_cap <- rbindlist(list(sol_2025, sol_2050, sol_2070))
# update tech
ESOL_2025@data$ESOL_2025_CL01 <- update(
  ESOL_2025@data$ESOL_2020_CL01,
  name = "ESOL_2025_CL01",
  desc = "Updated solar capacity, 2025, cluster 1",
  capacity = data.frame(
    region = sol_cap$reg7,
    year = sol_cap$year,
    stock = round(sol_cap$value / 1e3 / 3, 1) # drop crumbs, split equally to 3 clusters
  )
)
ESOL_2025@data$ESOL_2025_CL02 <- update(
  ESOL_2025@data$ESOL_2020_CL02,
  name = "ESOL_2025_CL02",
  desc = "Updated solar capacity, 2025, cluster 2",
  capacity = data.frame(
    region = sol_cap$reg7,
    year = sol_cap$year,
    stock = round(sol_cap$value / 1e3 / 3, 1) # drop crumbs, split equally to 3 clusters
  )
)
ESOL_2025@data$ESOL_2025_CL03 <- update(
  ESOL_2025@data$ESOL_2020_CL03,
  name = "ESOL_2025_CL03",
  desc = "Updated solar capacity, 2025, cluster 3",
  capacity = data.frame(
    region = sol_cap$reg7,
    year = sol_cap$year,
    stock = round(sol_cap$value / 1e3 / 3, 1) # drop crumbs, split equally to 3 clusters
  )
)
ESOL_2025@data$ESOL_2025_CL01@capacity
ESOL_2025@data$ESOL_2020_CL01 <- NULL # replaced with 2025
ESOL_2025@data$ESOL_2025_CL02@capacity
ESOL_2025@data$ESOL_2020_CL02 <- NULL # replaced with 2025
ESOL_2025@data$ESOL_2025_CL03@capacity
ESOL_2025@data$ESOL_2020_CL03 <- NULL # replaced with 2025

# replace "2020" with "2025" in technology names in for.sum
ESOL_2025@data$CT_ESOL_CL01@lhs[[1]]@for.sum$tech <-
  gsub("_2020_", "_2025_", 
       ESOL_2025@data$CT_ESOL_CL01@lhs[[1]]@for.sum$tech)

ESOL_2025@data$CT_ESOL_CL02@lhs[[1]]@for.sum$tech <-
  gsub("_2020_", "_2025_", 
       ESOL_2025@data$CT_ESOL_CL02@lhs[[1]]@for.sum$tech)

ESOL_2025@data$CT_ESOL_CL03@lhs[[1]]@for.sum$tech <-
  gsub("_2020_", "_2025_", 
       ESOL_2025@data$CT_ESOL_CL03@lhs[[1]]@for.sum$tech)
```

#### Wind
```{r}
elc$EWIN_w_const |> names()
EWIN_2025 <- elc$EWIN_w_const
EWIN_2025@data$CT_EWIN_CL01@rhs # check constraint

EWIN_2025@data$EWIN_2020_CL01@capacity # existing capacity
EWIN_2025@data$EWIN_2020_CL02@capacity # existing capacity
EWIN_2025@data$EWIN_2020_CL03@capacity # existing capacity

win_2025 <- new_capacity |> filter(grepl("Wind", source, ignore.case = T))
win_2025

win_2050 <- win_2025 |>
  mutate(
    year = 2050,
    value = 1 * value) # no retirement assumed

win_2070 <- win_2025 |>
  mutate(
    year = 2070,
    value = 0 * value) # full retirement assumed

win_cap <- rbindlist(list(win_2025, win_2050, win_2070))

# assign all to cluster 1 for simplicity
EWIN_2025@data$EWIN_2025_CL01 <- update(
  EWIN_2025@data$EWIN_2020_CL01,
  name = "EWIN_2020_CL01", # keep the name for constraint consistency
  desc = "Updated wind capacity, 2025, cluster 1",
  capacity = data.frame(
    region = win_cap$reg7,
    year = win_cap$year,
    stock = round(win_cap$value / 1e3, 1) # drop crumbs
  )
)
EWIN_2025@data$EWIN_2025_CL01@capacity
EWIN_2025@data$EWIN_2020_CL01 <- NULL # replaced with 2025

```

## Model configuration
```{r}
# # Update horizon up to 2075
# horizon_2020_2070_by_10 <-
#   newHorizon(
#   name = "Y2020_2070_by_10",
#   desc = "2020 to 2070 by 10 years",
#   period = 2020:2075,
#   intervals = c(1, 5, rep(10, 15)),
#   mid_is_end = F
# )
# horizon_2020_2070_by_10
# # (optional) Change milestone year 2023 to 2025
# horizon_2020_2070_by_10@intervals # see 2nd row
# horizon_2020_2070_by_10@intervals$mid[2] <- 2025
# horizon_2020_2070_by_10@intervals # check

# horizon_2020_2070_by_20 <-
#   newHorizon(
#   name = "Y2020_2070_by_20",
#   desc = "2020 to 2070 by 20 years",
#   period = 2020:2070,
#   intervals = c(1, 5, 10, rep(20, 15)),
#   mid_is_end = T
# )
# horizon_2020_2070_by_20@intervals
# $mid[3] <- 2050

horizon_2020_2070_by_15 <- newHorizon(
  name = "Y2020_2070_by_15",
  desc = "2020 to 2070 by 20 years",
  period = 2020:2070,
  intervals = c(1, 5, 10, rep(15, 5)),
  mid_is_end = T
)

# Updating constraints (to 2070+)
NO_NEW_HYDRO <- elc$NO_NEW_HYDRO
NO_NEW_HYDRO@rhs <- data.frame(
  year = 2020:2100, # extend the constraint
  rhs = 1e-20 # close to zero (non-zero to avoid dropping)
  ) 
NO_NEW_HYDRO@for.each <- data.frame(year = 2020:2100)
NO_NEW_HYDRO@lhs[[1]]@for.sum$tech <- 
  gsub("_2020", "_2025", 
       NO_NEW_HYDRO@lhs[[1]]@for.sum$tech)

NO_NEW_NUCLEAR <- elc$NO_NEW_NUCLEAR
NO_NEW_NUCLEAR@rhs <- data.frame(
  year = 2020:2100, # all range of years
  rhs = 1e-20
  ) 
NO_NEW_NUCLEAR@for.each <- data.frame(year = 2020:2100)
NO_NEW_NUCLEAR@lhs[[1]]@for.sum$tech <- 
  gsub("_2020", "_2025", 
       NO_NEW_NUCLEAR@lhs[[1]]@for.sum$tech)

# A model with the default configuration
mod <- newModel(
  name = glue("IDEEA_ELC_{regN}"),
  desc = "IDEEA electricity example model",
  region = unique(ideea_sf[[regN]]),
  discount = 0.05,
  calendar = ideea_modules$calendars$calendar_d365_h24,
  horizon = ideea_modules$horizons$horizon_2020_2060_by_10,
  data = newRepository(
    name = glue("repo_electricity_{regN}"),
    elc$repo_comm, # declaration of commodities used in the model
    # supply & import
    elc$repo_supply, # primary energy supply (markets)
    elc$repo_geoccs, # CCS storage potential by region
    # Generating technologies
    # elc$ECOA, # coal-fired generators
    ECOA_2025, # coal-fired generators updated to 2025
    # elc$EGAS, # gas-fired generators
    EGAS_2025, # gas-fired generators updated to 2025
    # elc$ENUC, # nuclear generators
    ENUC_2025, # nuclear generators updated to 2025
    # elc$EHYD, # hydro generators
    EHYD_2025, # hydro generators updated to 2025
    # elc$ESOL_w_const, # solar generators with constraints on capacity
    ESOL_2025, # solar generators updated to 2025 with constraints on capacity
    # elc$ESOL_w_land, # solar generators with constraints on land
    # elc$EWIN_w_const, # wind generators with constraints on capacity
    EWIN_2025, # wind generators updated to 2025 with constraints on capacity
    # elc$EWIN_w_land, # wind generators with constraints on land
    elc$EWIF_w_const, # wind generators with constraints on capacity
    # elc$EWIF_w_land, # wind generators with constraints on used surface area
    # elc$EBIO, # biomass generators
    EBIO_2025, # biomass generators updated to 2025
    elc$repo_ccstechs, # CCS technologies
    # capacity (weather) factors
    elc$WSOL, # solar capacity factor
    elc$WWIN, # onshore wind capacity 
    elc$WWIF, # offshore wind capacity 
    elc$WHYD, # hydro capacity factor
    # battery
    elc$STG_BTR, # battery storage with yearly cycle
    # STG_BTR_daily = STG_BTR_daily, # battery storage with daily cycle
    # transmission
    elc$repo_transmission_ac, # HVAC transmission network
    elc$repo_transmission_dc, # HVDC transmission network with converters
    # assumptions / policies
    NO_NEW_HYDRO, # no new hydro capacity
    NO_NEW_NUCLEAR, # no new nuclear capacity
    # elc$NO_NEW_WIND, # no new wind capacity
    # elc$NO_NEW_SOLAR, # no new solar capacity
    # demand
    # elc$DEMELC_BY, # electricity demand in the base year
    # elc$DEMELC_2X # electricity demand growth
    DEMELC_BY_2024, # electricity demand in the base year updated to 2024
    DEMELC_2X_2024 # electricity demand growth updated to 2024
  )
)
## Updating carbon policy to net zero by 2070
CO2_CAP_2070 <- ideea_modules$electricity[[regN]]$CO2_CAP
CO2_CAP_2070@rhs
CO2_CAP_2070@rhs$year[CO2_CAP_2070@rhs$year == 2060] <- 2070
CO2_CAP_2070@rhs

```

## Input check

### Wind clusters
<!-- ToDo: done -->
<!-- - add cluster information (N and tol) to the repository to avoid ambiguity   -->
<!-- - write a function for quick plotting of clustered data (wind/solar) - DONE -->

```{r}
names(elc$EWIN_w_const@data)
message("Number of regions: ", nreg)
```

### Plot wind capacity factor by cluster
<!-- ToDO: done -->
<!-- - debug gif function -->
<!-- - share IDEEA extra-repository -->
```{r}
# resource <- "win"; 
win_tol <- 0.05
calendar_1day_per_month <- 
  ideea_modules$calendars$calendar_d365_h24_subset_1day_per_month
  
slices_1day_per_month <- calendar_1day_per_month@timetable$slice

# ideea_sf <- get_ideea_map(nreg = nreg, offshore = T, islands = T)
ideea_win_cl_sf <- get_ideea_cl_sf(resource = "win", tol = win_tol)
plot(ideea_win_cl_sf["wcf_100m"])

win_cf_data <- get_ideea_cf("win", tol = win_tol, year = 2019)
win_cf_data
cf_name <- "wcf_100m"

# plot
wind_snapshot <- ideea_snapshot_cf(win_cf_data, ideea_win_cl_sf, ideea_sf, 
                  cf_name = cf_name, 
                  slice = slices_1day_per_month[100])
wind_snapshot
ggsave("images/wind_snapshot.png", wind_snapshot, scale = 1.5,
       width = 5, height = 5)

# gif
# ?ideea_gif_cf
ideea_gif_cf(x = win_cf_data, 
             ideea_cl_sf = ideea_win_cl_sf,
             ideea_sf = ideea_sf,
             cf_name = cf_name,
             slice = slices_1day_per_month[1:24],
             fps = 5, gif.width = 864, gif.height = 864,
             # fps = 5, gif.width = 432, gif.height = 432,
             filename = glue("wind_capacity_factors_sample.gif"))
```

### Plot solar capacity factors by cluster
```{r}
sol_tol <- 0.01
ideea_sol_cl_sf <- get_ideea_cl_sf(resource = "sol", tol = sol_tol)
sol_cf_data <- get_ideea_cf("sol", tol = sol_tol, year = 2019)
sol_cf_data
# plot
solar_snapshot <- ideea_snapshot_cf(sol_cf_data, ideea_sol_cl_sf, ideea_sf, 
                  cf_name = "scf_tl")
solar_snapshot
ggsave("images/solar_snapshot.png", solar_snapshot, scale = 1.5,
       width = 5, height = 5)
library(patchwork)
sol_win <- solar_snapshot + wind_snapshot
ggsave("images/solar_wind_snapshot.png", sol_win, scale = 1.5,
       width = 8.5, height = 5)

# gif
# sol_cf_data_EAST <- filter(sol_cf_data, reg5 == "EAST")
ideea_gif_cf(
  x = sol_cf_data,
  ideea_cl_sf = ideea_sol_cl_sf,
  ideea_sf = ideea_sf,
  cf_name = "scf_tl",
  slice = slices_1day_per_month[1:24],
  fps = 5, gif.width = 864, gif.height = 864,
  # fps = 5, gif.width = 332, gif.height = 432,
  filename = glue("sol_capacity_factors_sample_EAST.gif"))

```


### (optional) cluster locations
```{r, eval=FALSE}
fig.wind <- ggplot() +
  geom_sf(data = ideea_sf$geometry) +
  geom_sf(data = ideea_win_cl_sf, aes(fill = wcf_100m), color = "grey15") +
  scale_fill_viridis_c(option = "D", name = "CF") +
  labs(title = "Wind, 100m hub") +
  theme_ideea_map() +
  theme(plot.title = element_text(hjust = 0.5))

fig.sol <- ggplot() +
  geom_sf(data = ideea_sf$geometry) +
  geom_sf(data = ideea_sol_cl_sf, aes(fill = scf_tl), color = "grey85") +
  scale_fill_viridis_c(option = "A", name = "CF") +
  labs(title = "Solar, 1-axis tracking") +
  theme_ideea_map() +
  theme(plot.title = element_text(hjust = 0.5))

library(patchwork)
sol_win <- fig.sol + fig.wind
ggsave("images/solar_wind_cf.png", sol_win, scale = 1.5,
       width = 8.5, height = 5)

```


```{r}

EXPELC <- newExport(
  name = "EXPELC",
  desc = "Pseudo-export to capture curtailed generation",
  commodity = "ELC",
  unit = "cr.₹/GWh",
  exp = list(
    price = 0.001 # "cr.INR/GWh" 
    # convert("cr.₹/GWh", "cents/kWh", .001) # Check
    # convert("cr.₹/GWh", "₹/kWh", .001) # Check
  )
)

IMPELC <- newImport(
  name = "IMPELC",
  desc = "Unserved load penalty",
  commodity = "ELC",
  unit = "cr.₹/GWh",
  imp = list(
    price = 1e4 # high marginal price (penalty) for unserved load
    # convert("USD/kWh", "MUSD/GWh", 1e4)
  )
)
```



## Scenarios
Let's consider two scenarios for the electricity sector in 2050. The first scenario is a reference scenario (**REF**) without any policy intervention. The second scenario is a scenario with a carbon emissions constraint (**CAP**), targeting zero emissions by 2060 (linear reduction from 2030).  

Here we solve both scenarios for 2050 with full 8760 hours resolution and a subset of 1 day per month, 288 hours (24 * 12) in total. The subset scenario is relatively fast to solve and can be used for drafting and testing purposes. Though it does not provide a full picture of intermittent renewable generation and can downplay the role of storage and transmission.


### Reference scenario
The **REF** scenario is a business-as-usual that includes all the existing technologies that are already in the model. 

#### Solving for 288 hours
The 288 hours scenario can be solved with open source solvers like HiGHS or Cbc in minutes. Here we use the HiGHS solver with the barrier method and Julia interface.
```{r, eval=FALSE}
# (optional) set default solver
# set_default_solver(solver_options$julia_highs_parallel)
# set_progress_bar(type = "progress") # output progress bar in the console

# Sub-setting and interpolating of the model parameters
scen_REF_sub <- interpolate_model(
  mod,
  name = "REF25_sub", # indicating 'subset'
  calendar = ideea_modules$calendars$calendar_d365_h24_subset_1day_per_month,
  horizon = ideea_modules$horizons$horizon_2070
)
summary(scen_REF_sub)

# Writing the model script and data
scen_REF_sub <- write_sc(
  scen_REF_sub
  # solver = solver_options$julia_cplex
  # solver = solver_options$julia_highs_barrier
  # solver = solver_options$julia_highs
)
scen_REF_sub@status # info
scen_REF_sub@misc$tmp.dir # temporary directory

# Solving the scenario
scen_REF_sub <- solve_scenario(scen_REF_sub, wait = T)

# Reading the results
scen_REF_sub <- read(scen_REF_sub)
summary(scen_REF_sub)

# (optional) saving the scenario on disk
scen_REF_sub <- save_scenario(scen_REF_sub)
size(scen_REF_sub)

```

### Carbon emissions constraint scenario
The **CAP** scenario is a scenario with a carbon emissions constraint. Here we use pre-defined CO2 emissions constraint on the national level, assuming emissions trading. The constraint is saved in `ideea_modules[[regN]]$CO2_CAP` object.

#### Solving for 288 hours
```{r, eval=FALSE}
# Policy constraint 
# CO2_CAP <- ideea_modules$electricity[[regN]]$CO2_CAP

# Sub-setting and interpolating of the model parameters
scen_CAP_sub <- interpolate(
  mod,
  name = "CAP25_sub", # indicating 'subset'
  CO2_CAP_2070, # carbon emissions constraint
  calendar = ideea_modules$calendars$calendar_d365_h24_subset_1day_per_month,
  horizon = ideea_modules$horizons$horizon_2070
)

# Writing the model script and data
scen_CAP_sub <- write_sc(
  scen_CAP_sub
  # solver = solver_options$gams_gdx_cplex_parallel
  # solver = solver_options$julia_highs_barrier
  # solver = julia_highs_crossover_on
)

# Solving the scenario
scen_CAP_sub <- solve_scenario(scen_CAP_sub, wait = T)

# Reading the results
scen_CAP_sub <- read(scen_CAP_sub)
summary(scen_CAP_sub)

# (optional) saving the scenario on disk
scen_CAP_sub <- save_scenario(scen_CAP_sub)

```


## Solving for 8760 hours
The full-year scenario requires more computational resources and time. The generation of the problem also varies by the software. Here we use Python/Pyomo and CPLEX solver. The one-year scenario can be solved within an hour on a consumer-grade desktop or laptop. Barrier algorithm benefits from parallel processing and is normally faster on multi-core CPUs.  

### REF
```{r}
# Interpolating the model parameters
scen_REF <- interpolate(
  mod,
  name = "REF25",
  horizon = horizon_2020_2070_by_15
)

# Writing the model script and data
scen_REF <- write_sc(scen_REF)

# Solving the scenario
# scen_REF <- solve_scenario(scen_REF, wait = F)
save(scen_REF, file = file.path(ideea_scenarios(), "scen_REF.RData"))
rm(scen_REF); gc()
stop()
# 
# (load(file.path(ideea_scenarios(), "scen_REF.RData")))

# Reading the results
scen_REF <- read(scen_REF)
summary(scen_REF)

# (optional) saving the scenario on disk
scen_REF <- save_scenario(scen_REF)

```

### CAP
Similarly, the full-year scenario can be solved with the same solver and settings as **REF**.
```{r}
# Interpolating the model parameters
scen_CAP <- interpolate_model(
  mod, IMPELC, EXPELC,
  name = "CAP25",
  CO2_CAP_2070, # carbon emissions constraint
  horizon = horizon_2020_2070_by_15
)

# Writing the model script and data
scen_CAP <- write_script(scen_CAP)
  # replace default solver
  # solver_options$pyomo_cplex_barrier
  # solver_options$julia_cplex_barrier
  # solver = solver_options$gams_gdx_cplex_barrier
  # solver = solver_options$gams_gdx_cplex_parallel
  # solver = solver_options$julia_highs_barrier
  # solver = solver_options$julia_highs_parallel
# )

# Solving the scenario
# save(scen_CAP, file = file.path(ideea_scenarios(), "scen_CAP.RData"))
# rm(scen_CAP); gc()
scen_CAP <- solve_scenario(scen_CAP, wait = F)
# installr::os.sleep()
save(scen_CAP, file = file.path(ideea_scenarios(), "scen_CAP.RData"))
# stop()

# Reading the results
# (load(file.path(ideea_scenarios(), "scen_CAP.RData")))
scen_CAP <- read(scen_CAP)
summary(scen_CAP)
isInMemory(scen_CAP) 

# (optional) saving the scenario on disk
scen_CAP <- save_scenario(scen_CAP)
isInMemory(scen_CAP) # FALSE means the scenario is saved on disk
summary(scen_CAP) # check RAM size 
# installr::os.sleep()

```


## Results
... an example of processing some results ...

```{r}
all_slices <- scen_CAP@settings@calendar@timetable$slice

# (very) quick plots to use
?ideea_snapshot
ideea_snapshot(scen_REF, YEAR = 2050, SLICE = all_slices[1:24])
ideea_snapshot(scen_CAP, YEAR = 2050, SLICE = all_slices[1:24])
scen_cap_operation_data <- ideea_snapshot(scen_CAP, YEAR = 2050, 
               SLICE = all_slices, 
               return_data = T)
fwrite(scen_cap_operation_data, file = "tmp/scen_cap_operation_data.csv")

# ToDo: 
# - add functionality to plot several scenarios
# - fix colors / palette for fuel mix and techs

```



```{r}
# Combine scenarios in a named or unnamed list
# (in the case of unnamed list, scenario names are used)
sns <- list(scen_REF_sub, scen_CAP_sub)
# sns <- list(scen_REF_sub, scen_REF, scen_CAP_sub, scen_CAP)
# sns <- scen_REF_sub

```

Use `getData()` function to retrieve variables and parameters from solved scenario objects. See `?getData` for more details.
```{r}
# Objective function value (total discounted cost)
getData(sns, "vObjective", merge = T)
```
The "subset" time-slices scenarios normally show lower objective. 

Compare technological capacity by scenario.
```{r}
vTechCap <- getData(sns, "vTechCap", merge = T, process = T, 
                    tech_ = "^E",
                    digits = 2, drop.zeros = T) |> 
  drop_process_cluster() |> # removes cluster indicators from process names
  drop_process_vintage() |> # removes vintage indicators from process names
  group_by(scenario, process, year) |> # group and aggregate for figures/tables
  summarise(GW = sum(value), .groups = "drop") |>
  filter(!grepl("EDC", process)) |>
  as.data.table()
vTechCap |> pivot_wider(names_from = process, values_from = GW)

ggplot(vTechCap) +
  geom_bar(aes(x = scenario, y = GW, fill = process), stat = "identity") +
  scale_fill_viridis_d(option = "H", direction = -1) +
  theme_bw() 

vEmsFuelTot <- getData(sns, "vEmsFuelTot", merge = T) 

ggplot(vEmsFuelTot) +
  geom_bar(aes(x = scenario, y = value), stat = "identity", fill = "brown") +
  # scale_fill_viridis_d(option = "H", direction = -1) +
  theme_bw() 

```

