---
title: "IDEEA: LDV Passenger Vehicles"
editor_options:
  chunk_output_type: console
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.height = 4, fig.width = 7)
```

```{r libraries, message=FALSE, warning=FALSE}
# library(IDEEA)
# library(data.table)
source("R/setup.R")
source("R/ideea_report.r")
source("R/ideea_vehicle.R")
source("R/ideea_levcost.R")
source("R/read_transport_inputs.R")
```

## Read transport inputs

```{r read_inputs}
xls_path <- "data/transport_inputs.xlsx"
tabs     <- read_transport_inputs(xls_path)
names(tabs)
```

## Commodities

All commodities (fuels, transport services, infrastructure) are read from the
`commodities` tab. Individual objects are also extracted for direct use below.

```{r build_commodities}
repo_comm <- build_ldv_commodities(tabs)

# Emissions commodities from IDEEA modules (not in the Excel)
repo_comm <- add(repo_comm,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)

names(repo_comm)

```

## Vehicle parameters

```{r tech_params_table}
params_df <- tabs[["technologies"]]
knitr::kable(
  params_df[, c("name", "desc", "fuels", "eff_hwy", "eff_cty", "eff_unit",
                "invcost_2025", "fixom_2025", "olife")],
  caption = "LDV Passenger Vehicle Parameters (EPS-sourced: BNVFE eff, BNVP cost, AVMC O&M)"
)
```

## Build technology objects

```{r build_techs, results='hide'}
# Energy content (MJ/L or MJ/kg) by fuel type.
# For eff_unit = "MJ/km" (CNG/LPG/H2) the vector is not used for conversion
# but is retained for levcost calculations.
ldv_fuel_energy <- c(
  GSL   = 34.2,   # MJ/L LHV
  DSL   = 36.9,
  BIO   = 26.8,   # ethanol/biodiesel average
  LPG   = 26.0,
  CNG   = 35.8,   # MJ/m³ STP
  H2    = 120.0,  # MJ/kg LHV
  ELCPJ = NA_real_
)

repo_tech <- build_ldv_technologies(tabs, fuel_energy = ldv_fuel_energy)
repo_tech <- build_ldv_fleet(repo_tech, tabs)
names(repo_tech)
```

Number of technologies created: `r length(repo_tech@data)`.

## Technology diagrams

```{r draw_techs, fig.height=4, fig.width=9, results='asis'}
for (nm in names(repo_tech@data)) {
  cat("\n\n###", nm, "\n\n")
  # draw(repo_tech@data[[nm]])
}
```

## Fuel efficiency summary

```{r eff_summary, results='asis'}
eff_tables <- lapply(names(repo_tech@data), function(nm) {
  df <- vehicle_eff_table(repo_tech@data[[nm]], fuel_energy = ldv_fuel_energy)
  if (!is.null(df)) { df$tech <- nm; df }
})
eff_all <- do.call(rbind, Filter(Negate(is.null), eff_tables))

knitr::kable(
  eff_all[, c("tech", "mode", "fuel", "L_per_100km", "km_per_kWh", "MJ_per_km", "mpg_us")],
  digits  = 2,
  caption = "Fuel efficiency summary (all units)"
)
```





## Levelized cost repository

Supply objects with representative fuel costs (MUSD/PJ) used when computing
levelized costs inside `ideea_report()`. Technologies that don't consume a given
fuel simply ignore the corresponding supply.  `ideea_levcost()` auto-creates
zero-cost supply for any input not covered here.

```{r levcost_repo}
repo_sup <- build_ldv_supply(tabs)

# Levelized-cost repository: commodities + supply objects + emissions
levcost_repo <- Reduce(add, c(
  unname(repo_comm@data),
  unname(repo_sup@data)
), newRepository("levcost_repo"))
```

## Generate individual PDF reports

Reports are written to `reports/LDV/`. Each report is one page (PDF).
Levelized costs are computed automatically via `ideea_levcost()` and embedded
as a compact NPV-breakdown figure (and production frontier when applicable).

```{r generate_reports, results='asis'}
dir.create("reports/LDV", recursive = TRUE, showWarnings = FALSE)

# Tech name → image file mapping (edit as needed)
ldv_images <- c(
  LDV_GSL  = "images/techs/LDV-gasoline.png",
  LDV_DSL  = "images/techs/LDV-diesel.png",
  LDV_CNG  = "images/techs/LDV-CNG.png",
  LDV_LPG  = "images/techs/LDV-PLG.png",
  LDV_HYB  = "images/techs/LDV-hybrid.png",
  LDV_PHEV = "images/techs/LDV-PHEV.png",
  LDV_EV   = "images/techs/LDV-electric.png",
  LDV_H2   = "images/techs/LDV-hydrogen.png",
  LDV_BIO  = "images/techs/LDV-biofuel.png",
  LDV_E85  = "images/techs/LDV-E85.png",
  LDV_E100 = "images/techs/LDV-E100.png",
  LDV_B100 = "images/techs/LDV-B100.png"
)

for (nm in names(repo_tech@data)) {
  tch <- repo_tech@data[[nm]]
  img <- if (nm %in% names(ldv_images) && file.exists(ldv_images[[nm]]))
    ldv_images[[nm]] else NULL
  
  if (F) {
    # check levcost
    lc <- ideea_levcost(
      object = tch,
      group = "o",
      repo = levcost_repo,
      discount    = 0.07, 
      base_year   = 2024
      )
    
    # plots
    # per-year stacked components + NPV line
    print(autoplot(lc))
    # NPV-weighted breakdown (or by-comm if frontier)    
    print(autoplot(lc, type = "npv"))
    if (!is.null(lc$frontier)) {
      autoplot(lc, type = "frontier") +
        scale_x_continuous(limits = c(0, 1)) +
        scale_y_continuous(limits = c(0, 1))
    }
  }

  tryCatch({
    ideea_report(
      object      = tch,
      type        = "vehicle",
      image_file  = img,
      file        = file.path("reports/LDV", nm),
      format      = c("pdf", "html"),
      fuel_energy = ldv_fuel_energy,
      # levcost args – picked up by ideea_report() and forwarded to ideea_levcost()
      group       = "o",
      repo        = levcost_repo@data,
      discount    = 0.07,
      base_year   = 2024,
      verbose     = FALSE
    )
    cat("- **", nm, "** → `reports/LDV/", nm, ".pdf` + `.html`\n", sep = "")
  }, error = function(e) {
    cat("- **", nm, "** — ERROR:", conditionMessage(e), "\n")
  })
}
```
