---
title: "LDV Technical and Economic Parameters"
subtitle: "Indicative ranges for the USA and India"
author: "(LLM assembled)"
# date: last-modified
format:
  revealjs:
    theme: [default, styles.scss]
    slide-number: true
    progress: true
    controls: true
    transition: fade
    transition-speed: fast
    center: false
    margin: 0.06
    highlight-style: github
    code-line-numbers: false
footer: "WRI · IDEEA"
logo: images/logo_wri.png
execute:
  echo: false
  warning: false
  message: false
params:
  currency: "both"       # "USD", "INR", or "both"
  usd_to_inr: 84        # exchange rate; override with -P usd_to_inr:85
  fig_base_currency: "INR"  # when currency=="both": "INR" (primary) or "USD" (primary)
editor_options:
  chunk_output_type: console
---

```{r setup}
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(scales)

# ── Currency helpers ──────────────────────────────────────────────────────────
.cur       <- params$currency           # "USD", "INR", or "both"
.rate      <- params$usd_to_inr         # e.g. 84
show_both  <- .cur == "both"
cur_factor <- if (.cur == "INR") .rate else 1   # multiplier for cost values
cur_sym    <- if (.cur == "INR") "\u20b9" else "$"  # ₹ or $

# Figure-specific helpers: primary axis currency when currency == "both"
.fig_base  <- params$fig_base_currency              # "INR" or "USD"
fig_factor <- if (show_both) (if (.fig_base == "INR") .rate else 1) else cur_factor

# Label for figure primary axis
fig_lbl <- function(unit) {
  if (!show_both) return(cur_lbl(unit))
  if (.fig_base == "INR") gsub("USD", "\u20b9", unit) else unit
}

# Secondary (reference) axis added to figures when currency == "both"
ref_axis <- function(primary_label) {
  if (!show_both) return(waiver())
  if (.fig_base == "INR") {
    sec_axis(~ . / .rate, name = gsub("\u20b9", "USD", primary_label))
  } else {
    sec_axis(~ . * .rate, name = gsub("USD", "\u20b9", primary_label))
  }
}

# Replace "USD" with "INR" (or "USD / INR") in unit labels
cur_lbl <- function(unit) {
  if (show_both) {
    gsub("USD", paste0("USD / \u20b9INR (1 USD = ", .rate, " \u20b9)"), unit)
  } else if (.cur == "INR") {
    gsub("USD", "\u20b9", unit)
  } else {
    unit
  }
}

# Format a numeric range with optional unit; applies currency scaling
fmt_cur <- function(lo, hi, unit = "", scale = 1) {
  if (show_both) {
    usd  <- fmt_range(lo * scale, hi * scale, paste("USD", unit))
    inr  <- fmt_range(lo * scale * .rate, hi * scale * .rate, paste0("\u20b9 ", unit))
    paste0(usd, " / ", inr)
  } else {
    fmt_range(lo * scale * cur_factor, hi * scale * cur_factor, cur_lbl(unit))
  }
}

fmt_range <- function(lo, hi, unit = "") {
  ifelse(
    is.na(lo) | is.na(hi),
    "n/a",
    paste0(
      format(lo, trim = TRUE, big.mark = ","),
      "\u2013",
      format(hi, trim = TRUE, big.mark = ","),
      ifelse(unit == "", "", paste0(" ", unit))
    )
  )
}

vehicle_order <- c(
  "Gasoline ICE", "Diesel ICE", "Hybrid gasoline HEV", "Plug-in hybrid PHEV",
  "CNG / LPG", "Battery electric BEV", "Hydrogen fuel cell FCEV"
)

# ── Technical data ────────────────────────────────────────────────────────────
tech <- tibble::tribble(
  ~vehicle,                    ~efficiency,                              ~ttw_efficiency,       ~life_km_lo, ~life_km_hi, ~weight_t_lo, ~weight_t_hi, ~passengers_lo, ~passengers_hi,
  "Gasoline ICE",              "6\u20139 L/100 km",                      "20\u201325%",         150000, 250000, 1.2, 1.8, 1.5, 2.0,
  "Diesel ICE",                "4\u20137 L/100 km",                      "25\u201330%",         200000, 300000, 1.3, 2.0, 1.5, 2.0,
  "Hybrid gasoline HEV",       "3.5\u20135.5 L/100 km",                  "30\u201335%",         200000, 300000, 1.4, 1.9, 1.5, 2.0,
  "Plug-in hybrid PHEV",       "2\u20134 L/100 km + 15\u201325 kWh/100 km", "35\u201345% blended", 200000, 300000, 1.6, 2.2, 1.5, 2.0,
  "CNG / LPG",                 "5\u20138 kg CNG-eq./100 km",             "25\u201330%",         200000, 300000, 1.3, 2.0, 1.5, 2.0,
  "Battery electric BEV",      "13\u201320 kWh/100 km",                  "75\u201390%",         200000, 300000, 1.6, 2.5, 1.5, 2.0,
  "Hydrogen fuel cell FCEV",   "0.8\u20131.2 kg H2/100 km",              "50\u201360%",         200000, 300000, 1.8, 2.5, 1.5, 2.0
) %>%
  mutate(vehicle = factor(vehicle, levels = vehicle_order))

# ── Economic data (values in USD) ─────────────────────────────────────────────
capex <- tibble::tribble(
  ~vehicle,                    ~country, ~capex_lo, ~capex_hi,
  "Gasoline ICE",              "USA",    22000,  35000,
  "Gasoline ICE",              "India",   6000,  15000,
  "Diesel ICE",                "USA",    25000,  40000,
  "Diesel ICE",                "India",   8000,  18000,
  "Hybrid gasoline HEV",       "USA",    28000,  40000,
  "Hybrid gasoline HEV",       "India",  15000,  30000,
  "Plug-in hybrid PHEV",       "USA",    35000,  50000,
  "Plug-in hybrid PHEV",       "India",  25000,  45000,
  "CNG / LPG",                 "USA",    24000,  35000,
  "CNG / LPG",                 "India",   7000,  16000,
  "Battery electric BEV",      "USA",    30000,  55000,
  "Battery electric BEV",      "India",  12000,  35000,
  "Hydrogen fuel cell FCEV",   "USA",    50000,  70000,
  "Hydrogen fuel cell FCEV",   "India",     NA,     NA
) %>% mutate(vehicle = factor(vehicle, levels = vehicle_order))

fom <- tibble::tribble(
  ~vehicle,                    ~country, ~fom_lo, ~fom_hi,
  "Gasoline ICE",              "USA",     800,  1500,
  "Gasoline ICE",              "India",   200,   600,
  "Diesel ICE",                "USA",     800,  1500,
  "Diesel ICE",                "India",   200,   600,
  "Hybrid gasoline HEV",       "USA",     700,  1300,
  "Hybrid gasoline HEV",       "India",   250,   700,
  "Plug-in hybrid PHEV",       "USA",     800,  1400,
  "Plug-in hybrid PHEV",       "India",   300,   800,
  "CNG / LPG",                 "USA",     800,  1300,
  "CNG / LPG",                 "India",   250,   700,
  "Battery electric BEV",      "USA",     400,   900,
  "Battery electric BEV",      "India",   150,   500,
  "Hydrogen fuel cell FCEV",   "USA",    1000,  2000,
  "Hydrogen fuel cell FCEV",   "India",    NA,    NA
) %>% mutate(vehicle = factor(vehicle, levels = vehicle_order))

varcost <- tibble::tribble(
  ~vehicle,                    ~country, ~var_lo, ~var_hi,
  "Gasoline ICE",              "USA",    0.07, 0.12,
  "Gasoline ICE",              "India",  0.08, 0.14,
  "Diesel ICE",                "USA",    0.06, 0.10,
  "Diesel ICE",                "India",  0.06, 0.11,
  "Hybrid gasoline HEV",       "USA",    0.05, 0.09,
  "Hybrid gasoline HEV",       "India",  0.06, 0.10,
  "Plug-in hybrid PHEV",       "USA",    0.04, 0.10,
  "Plug-in hybrid PHEV",       "India",  0.05, 0.09,
  "CNG / LPG",                 "USA",    0.05, 0.09,
  "CNG / LPG",                 "India",  0.04, 0.07,
  "Battery electric BEV",      "USA",    0.02, 0.05,
  "Battery electric BEV",      "India",  0.01, 0.03,
  "Hydrogen fuel cell FCEV",   "USA",    0.10, 0.20,
  "Hydrogen fuel cell FCEV",   "India",    NA,   NA
) %>% mutate(vehicle = factor(vehicle, levels = vehicle_order))

# Fuel prices — numeric (USD) to allow INR conversion
fuel_prices <- tibble::tribble(
  ~energy_carrier, ~country, ~price_lo, ~price_hi, ~unit,
  "Gasoline",      "USA",    0.80, 1.20, "/L",
  "Gasoline",      "India",  1.20, 1.50, "/L",
  "Diesel",        "USA",    0.90, 1.30, "/L",
  "Diesel",        "India",  1.00, 1.30, "/L",
  "Electricity",   "USA",    0.12, 0.25, "/kWh",
  "Electricity",   "India",  0.06, 0.12, "/kWh",
  "CNG",           "USA",    0.80, 1.50, "/kg",
  "CNG",           "India",  0.70, 1.00, "/kg",
  "Hydrogen",      "USA",   12.00, 20.00, "/kg",
  "Hydrogen",      "India",   NA,   NA,   "/kg"
)

# ── LCOM calculation ──────────────────────────────────────────────────────────
crf <- function(discount_rate, lifetime_years) {
  r <- discount_rate
  n <- lifetime_years
  r * (1 + r)^n / ((1 + r)^n - 1)
}

lcom_assumptions <- tibble::tribble(
  ~country, ~annual_km, ~lifetime_years, ~discount_rate,
  "USA",    15000,      12,              0.06,
  "India",  10000,      12,              0.10
) %>%
  mutate(crf = crf(discount_rate, lifetime_years))

lcom <- capex %>%
  left_join(fom,      by = c("vehicle", "country")) %>%
  left_join(varcost,  by = c("vehicle", "country")) %>%
  left_join(lcom_assumptions %>% select(country, annual_km, crf), by = "country") %>%
  mutate(
    capex_mid        = (capex_lo + capex_hi) / 2,
    fom_mid          = (fom_lo   + fom_hi)   / 2,
    var_mid          = (var_lo   + var_hi)   / 2,
    capex_component  = capex_mid * crf / annual_km,
    fom_component    = fom_mid   / annual_km,
    variable_component = var_mid,
    lcom_mid = capex_component + fom_component + variable_component,
    lcom_lo  = capex_lo * crf / annual_km + fom_lo / annual_km + var_lo,
    lcom_hi  = capex_hi * crf / annual_km + fom_hi / annual_km + var_hi,
    vehicle  = factor(vehicle, levels = vehicle_order)
  )
```

## Scope {.smaller}

| Code | Technology |
|------|-----------|
| **Gasoline ICE** | Gasoline internal combustion engine vehicle |
| **Diesel ICE** | Diesel internal combustion engine vehicle |
| **HEV** | Hybrid electric vehicle (gasoline) |
| **PHEV** | Plug-in hybrid electric vehicle |
| **CNG / LPG** | Compressed natural gas / liquefied petroleum gas |
| **BEV** | Battery electric vehicle |
| **FCEV** | Hydrogen fuel-cell electric vehicle |

Countries covered: **USA** and **India**

---

## Technical parameters {.scrollable .smaller}

```{r tech-table}
tech_table <- tech %>%
  mutate(
    `Lifespan`           = fmt_range(life_km_lo / 1000, life_km_hi / 1000, "thousand km"),
    `Weight`             = fmt_range(weight_t_lo, weight_t_hi, "t"),
    `Passengers/vehicle` = fmt_range(passengers_lo, passengers_hi, "persons")
  ) %>%
  select(
    Vehicle                    = vehicle,
    `Energy efficiency`        = efficiency,
    `Tank-to-wheel efficiency` = ttw_efficiency,
    `Lifespan`,
    `Weight`,
    `Passengers/vehicle`
  )

kable(tech_table, caption = "Indicative technical parameters by LDV technology.") %>%
  kable_styling(font_size = 13, bootstrap_options = c("striped", "condensed"), full_width = TRUE)
```

---

## Vehicle lifetime

```{r lifespan-figure, fig.height=4.5, fig.width=8}
tech %>%
  ggplot(aes(y = vehicle)) +
  geom_linerange(aes(xmin = life_km_lo / 1000, xmax = life_km_hi / 1000),
                 linewidth = 5, colour = "#1f78b4") +
  labs(x = "Vehicle lifetime, thousand km", y = NULL,
       title = "Indicative lifetime range by technology") +
  theme_minimal(base_size = 13)
```

---

## Vehicle weight

```{r weight-figure, fig.height=4.5, fig.width=8}
tech %>%
  ggplot(aes(y = vehicle)) +
  geom_linerange(aes(xmin = weight_t_lo, xmax = weight_t_hi),
                 linewidth = 5, colour = "#33a02c") +
  labs(x = "Curb weight, tonnes", y = NULL,
       title = "Indicative weight range by technology") +
  theme_minimal(base_size = 13)
```

---

## Investment cost (CAPEX) {.scrollable .smaller}

```{r capex-table}
capex_tbl <- capex %>%
  mutate(`Investment cost` = fmt_cur(capex_lo / 1000, capex_hi / 1000,
                                     ifelse(.cur == "INR", "thousand \u20b9/vehicle",
                                            "thousand USD/vehicle"))) %>%
  select(Vehicle = vehicle, Country = country, `Investment cost`) %>%
  pivot_wider(names_from = Country, values_from = `Investment cost`)

kable(capex_tbl, caption = "Investment cost (thousand per vehicle)") %>%
  kable_styling(font_size = 12, bootstrap_options = c("striped", "condensed"), full_width = TRUE)
```

```{r capex-figure, fig.height=3, fig.width=9}
.xl <- paste("Investment cost, thousand", fig_lbl("USD"), "/ vehicle")
capex %>%
  filter(!is.na(capex_lo)) %>%
  mutate(lo = capex_lo * fig_factor / 1000,
         hi = capex_hi * fig_factor / 1000) %>%
  ggplot(aes(y = vehicle)) +
  geom_linerange(aes(xmin = lo, xmax = hi), linewidth = 4, colour = "#e31a1c") +
  facet_wrap(~country) +
  (if (show_both) scale_x_continuous(sec.axis = ref_axis(.xl)) else NULL) +
  labs(x = .xl, y = NULL) +
  theme_minimal(base_size = 10)
```

---

## Fixed O&M cost {.scrollable .smaller}

```{r fom-table}
fom_tbl <- fom %>%
  mutate(`Fixed O&M` = fmt_cur(fom_lo, fom_hi,
                               ifelse(.cur == "INR", "\u20b9/year", "USD/year"))) %>%
  select(Vehicle = vehicle, Country = country, `Fixed O&M`) %>%
  pivot_wider(names_from = Country, values_from = `Fixed O&M`)

kable(fom_tbl, caption = "Fixed annual O&M cost per vehicle") %>%
  kable_styling(font_size = 12, bootstrap_options = c("striped", "condensed"), full_width = TRUE)
```

```{r fom-figure, fig.height=3, fig.width=9}
.xl <- paste("Fixed O&M,", fig_lbl("USD"), "/ year")
fom %>%
  filter(!is.na(fom_lo)) %>%
  mutate(lo = fom_lo * fig_factor,
         hi = fom_hi * fig_factor) %>%
  ggplot(aes(y = vehicle)) +
  geom_linerange(aes(xmin = lo, xmax = hi), linewidth = 4, colour = "#ff7f00") +
  facet_wrap(~country) +
  (if (show_both) scale_x_continuous(sec.axis = ref_axis(.xl)) else NULL) +
  labs(x = .xl, y = NULL) +
  theme_minimal(base_size = 10)
```

---

## Fuel & energy prices {.smaller}

```{r fuel-price-table}
fp_display <- fuel_prices %>%
  rowwise() %>%
  mutate(
    `Price range` = if (is.na(price_lo)) {
      "not commercial"
    } else if (show_both) {
      paste0(
        fmt_range(price_lo, price_hi, paste0("USD", unit)),
        " / ",
        fmt_range(price_lo * .rate, price_hi * .rate, paste0("\u20b9", unit))
      )
    } else {
      fmt_range(price_lo * cur_factor, price_hi * cur_factor,
                paste0(cur_sym, unit))
    }
  ) %>%
  ungroup() %>%
  select(`Energy carrier` = energy_carrier, Country = country, `Price range`)

kable(fp_display,
      caption = paste0("Indicative fuel and energy price ranges",
                       if (.cur != "USD") paste0(" (1 USD = \u20b9", .rate, ")") else "")) %>%
  kable_styling(font_size = 13, bootstrap_options = c("striped", "condensed"), full_width = TRUE)
```

---

## Variable cost per km {.scrollable .smaller}

```{r varcost-table}
vc_tbl <- varcost %>%
  mutate(`Variable cost` = fmt_cur(var_lo, var_hi,
                                   ifelse(.cur == "INR", "\u20b9/km", "USD/km"))) %>%
  select(Vehicle = vehicle, Country = country, `Variable cost`) %>%
  pivot_wider(names_from = Country, values_from = `Variable cost`)

kable(vc_tbl, caption = "Fuel/energy variable cost per vehicle-km") %>%
  kable_styling(font_size = 12, bootstrap_options = c("striped", "condensed"), full_width = TRUE)
```

```{r varcost-figure, fig.height=3, fig.width=9}
.xl <- paste("Variable cost,", fig_lbl("USD"), "/ km")
varcost %>%
  filter(!is.na(var_lo)) %>%
  mutate(lo = var_lo * fig_factor,
         hi = var_hi * fig_factor) %>%
  ggplot(aes(y = vehicle)) +
  geom_linerange(aes(xmin = lo, xmax = hi), linewidth = 4, colour = "#6a3d9a") +
  facet_wrap(~country) +
  (if (show_both) scale_x_continuous(sec.axis = ref_axis(.xl)) else NULL) +
  labs(x = .xl, y = NULL) +
  theme_minimal(base_size = 10)
```

---

## Key modeling implications {.smaller}

**Dominant cost driver by drivetrain**

| Technology | Dominant driver |
|-----------|----------------|
| **ICE gasoline / diesel** | Fuel cost |
| **HEV / PHEV** | Fuel economy, electric driving share, annual mileage |
| **CNG / LPG** | Local fuel price and refueling-network access |
| **BEV** | Capital cost, battery lifetime, electricity tariff, annual mileage |
| **FCEV** | Hydrogen price (current market conditions) |

**Country-level sensitivities**

- **India** has higher discount rates (8–15 %) → capital-intensive technologies (BEV, PHEV) face greater cost penalty than in the USA
- BEV has a **strong electricity cost advantage** in India ($0.06–0.12/kWh vs $0.12–0.25/kWh in the USA)
- Vehicle-size bias: USA leans toward larger vehicles and SUVs; India toward smaller, lower-cost vehicles

---

## Country modeling assumptions {.smaller}

```{r country-assumptions}
country_assumptions <- tibble::tribble(
  ~Parameter,               ~USA,                         ~India,
  "Annual mileage",         "12\u201320 thousand km/year",  "8\u201315 thousand km/year",
  "Discount rate",          "5\u20138%",                    "8\u201315%",
  "Fuel cost pressure",     "Moderate",                   "High",
  "Electricity advantage",  "Moderate",                   "Strong",
  "Capital-cost sensitivity","Medium",                    "Very high",
  "Vehicle-size bias",      "Larger vehicles and SUVs",   "Smaller and lower-cost vehicles"
)

kable(country_assumptions,
      caption = "Suggested country-level assumptions for levelized cost analysis.") %>%
  kable_styling(font_size = 13, bootstrap_options = c("striped", "condensed"), full_width = TRUE)
```

Baseline used below: **USA** — 15 000 km/year, 12 years, 6 % discount rate; **India** — 10 000 km/year, 12 years, 10 % discount rate.

---
---

## LCOM results {.scrollable .smaller}

```{r lcom-results-table}
cur_unit <- if (.cur == "INR") "\u20b9/km" else "USD/km"

lcom_tbl <- lcom %>%
  filter(!is.na(lcom_mid)) %>%
  mutate(
    `LCOM range`      = fmt_cur(lcom_lo, lcom_hi, cur_unit),
    `Baseline`        = fmt_cur(lcom_mid, lcom_mid, cur_unit) %>%
                          gsub("\u2013.*", "", .),  # midpoint only
    `Capital`         = round(capex_component * cur_factor, 3),
    `Fixed O&M`       = round(fom_component   * cur_factor, 3),
    `Fuel / variable` = round(variable_component * cur_factor, 3)
  ) %>%
  select(Vehicle = vehicle, Country = country,
         `LCOM range`, `Baseline`,
         `Capital`, `Fixed O&M`, `Fuel / variable`) %>%
  arrange(Country, `Baseline`)

kable(lcom_tbl,
      caption = paste0("Illustrative total LCOM by vehicle type and country (",
                       cur_unit, ").")) %>%
  kable_styling(font_size = 12, bootstrap_options = c("striped", "condensed"), full_width = TRUE)
```

---

## LCOM comparison

```{r lcom-range-figure, fig.height=4.8, fig.width=9}
.xl <- paste("Levelized cost of mobility,", fig_lbl("USD"), "/ km")
lcom %>%
  filter(!is.na(lcom_mid)) %>%
  mutate(
    lo  = lcom_lo  * fig_factor,
    hi  = lcom_hi  * fig_factor,
    mid = lcom_mid * fig_factor
  ) %>%
  ggplot(aes(y = vehicle)) +
  geom_linerange(aes(xmin = lo, xmax = hi), linewidth = 4, colour = "#1f78b4", alpha = 0.6) +
  geom_point(aes(x = mid), size = 3, colour = "#e31a1c") +
  facet_wrap(~country) +
  (if (show_both) scale_x_continuous(sec.axis = ref_axis(.xl)) else NULL) +
  labs(x = .xl, y = NULL, title = "Total LCOM range (bar) and baseline midpoint (dot)") +
  theme_minimal(base_size = 12)
```

---

## LCOM cost decomposition

```{r lcom-decomp-figure, fig.height=4.8, fig.width=9}
.xl <- paste("Cost component,", fig_lbl("USD"), "/ km")
lcom %>%
  filter(!is.na(lcom_mid)) %>%
  select(vehicle, country,
         Capital         = capex_component,
         `Fixed O&M`     = fom_component,
         `Fuel / variable` = variable_component) %>%
  pivot_longer(cols = c(Capital, `Fixed O&M`, `Fuel / variable`),
               names_to = "component", values_to = "value") %>%
  mutate(value = value * fig_factor) %>%
  ggplot(aes(x = value, y = vehicle, fill = component)) +
  geom_col() +
  facet_wrap(~country) +
  scale_fill_brewer(palette = "Set1") +
  (if (show_both) scale_x_continuous(sec.axis = ref_axis(.xl)) else NULL) +
  labs(
    x     = .xl,
    y     = NULL,
    fill  = NULL,
    title = "Baseline LCOM decomposition by cost component"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

---

## Technology ranking {.smaller}

```{r lcom-ranking-table}
lcom_rank <- lcom %>%
  filter(!is.na(lcom_mid)) %>%
  group_by(country) %>%
  arrange(lcom_mid, .by_group = TRUE) %>%
  mutate(rank = row_number()) %>%
  ungroup() %>%
  mutate(
    `Baseline LCOM` = paste0(
      format(round(lcom_mid * cur_factor, 3), nsmall = 3),
      " ", if (.cur == "INR") "\u20b9" else "$", "/km"
    )
  ) %>%
  select(Country = country, Rank = rank, Vehicle = vehicle, `Baseline LCOM`)

kable(lcom_rank,
      caption = paste0("Technology ranking by baseline LCOM",
                       if (.cur != "USD") paste0(" (1 USD = \u20b9", .rate, ")") else "")) %>%
  kable_styling(font_size = 13, bootstrap_options = c("striped", "condensed"), full_width = TRUE)
```

::: {.callout-note appearance="minimal"}
Results are highly sensitive to annual mileage — higher utilisation favours capital-intensive BEV and PHEV.
:::


