---
title: "New Data: 2024-2025"
format: html
editor_options: 
  chunk_output_type: console
---

Near term ToDo list:

add state-level constraints for renewables potential from shared xls-worksheet

electric load figures by state/month (consider a function) add (finish) simplified/clustered transmission network from PyPSA-Earth model reconstruct levelized costs function (\~end of Oct.) new sectors (which one is the next - to discuss on the next meeting) IDEEA technical note (when the first version from Varun is available)

## Setup

```{r}
library(tidyverse)
library(data.table)
library(glue)
library(plotly)

# select number of regions
# nreg <- 5 # not implemented
nreg <- 32
TZ <- "Asia/Kolkata"
```

## Installed Capacity

```{r}
# Path to the data
# installed_capacity_r5 <- "data/Installed capacity/Region wise"
installed_capacity_r32 <- "data/Installed capacity/State wise"

# files 
ff <- list.files(get(glue("installed_capacity_r{nreg}")), pattern = "xlsx$", full.names = T)

xls_names <- readxl::excel_sheets(ff[1])

ll <- list()
for (i in seq_along(ff)) {
  f <- ff[i]
  message("Reading: '", basename(f), "'")
  ll[[i]] <- 
    readxl::read_excel(f, sheet = xls_names[1], na = c("", "-")) |>
    # transform into long format
    pivot_longer(cols = matches("^20"), 
                 names_to = c("year", "month"), 
                 names_pattern = "(.*)-(.*)",
                 names_transform = as.integer) |>
    # filter out ending-info-rows
    filter(!grepl("NITI|https", Parameter)) |>
    filter(Parameter != "") |>
    # optional filter out "Totals"
    filter(!grepl("Total", Parameter)) |>
    # Split "source" and "units"
    mutate(
      date = as.Date(paste0(year, "-", month, "-", "15"), tz = TZ), # assuming mid-month
      source = sub(" \\(.*", "", Parameter),
      unit = sub(".*\\(in ([^)]+)\\).*", "\\1", Parameter),
      .before = 1
    ) |>
    as.data.table()
}

# Merge the list into one table
gen_capacity_2025 <- rbindlist(ll) |> select(-Parameter) 
save(gen_capacity_2025, file = "data/gen_capacity_2025.RData")

```

### Plot capacity

```{r}
p <- ggplot(gen_capacity_2025) +
  geom_col(aes(date, value/1e3, fill = source)) +
  scale_fill_viridis_d(option = "H", name = "GW", direction = -1) +
  facet_wrap(~State, scale = "free_y") +
  theme_bw()
# p
ggsave("images/gen_capacity_2025.png", p, scale = 2, width = 10, height = 8)
ggplotly(p)

```

## Daily generation

```{r}
daily_generation_r32 <- "data/Daily Generation/State-wise"

# files 
ff <- list.files(daily_generation_r32, pattern = "xlsx$", full.names = T)
message("Found ", length(ff), " files")

xls_names <- readxl::excel_sheets(ff[1])
# f <- ff[1]

ll <- list()
for (i in seq_along(ff)) {
  f <- ff[i]
  message("Reading: '", basename(f), "'")
  ll[[i]] <-
    readxl::read_excel(f, sheet = xls_names[1], na = c("", "-")) |>
    # transform into long format
    pivot_longer(cols = matches("^20"), names_to = "Date"
                 # names_to = c("year", "month"), 
                 # names_pattern = "(.*)-(.*)",
                 # names_transform = as.Date
                 ) |>
      # as.data.table()
    # filter out ending-info-rows
    filter(!grepl("NITI|https", Parameter)) |>
    filter(Parameter != "") |>
    # optional filter out "Totals"
    filter(!grepl("Total", Parameter)) |>
    # Split "source" and "units"
    mutate(
      # date = as.Date(paste0(year, "-", month, "-", "15")), # assuming mid-month
      Date = as.Date(Date, tz = TZ),
      source = sub(" \\(.*", "", Parameter),
      source = sub(" - Generation", "", source) |> trimws(),
      unit = sub(".*\\(in ([^)]+)\\).*", "\\1", Parameter) |> trimws(),
      .before = 1
    ) |>
    as.data.table()
}

# Merge the list into one table
generation_daily_2025 <- rbindlist(ll) |> select(-Parameter) 
save(generation_daily_2025, file = "data/generation_daily_2025.RData")

```

### Plot generation

```{r}
p <- ggplot(generation_daily_2025) +
  geom_col(aes(Date, value, fill = source)) +
  scale_fill_viridis_d(option = "H", name = "MWh (MU)", direction = -1) +
  facet_wrap(~State, scale = "free_y") +
  theme_bw()
# p
ggsave("images/generation_daily_2025.png", p, scale = 1.5, width = 10, 
       height = 8)
ggplotly(p)

```

## Load curve

```{r}
# path to the data
load_curve_r32 <- "data/Load curve/State-wise"

# files 
ff <- list.files(load_curve_r32, pattern = "xlsx$", full.names = T, 
                 recursive = T)
message("Found ", length(ff), " files")

f <- ff[1]
xls_names <- readxl::excel_sheets(f)

ll <- list()
for (i in seq_along(ff)) {
  f <- ff[i]
  message("Reading: '", basename(f), "'")
  ll[[i]] <-
    readxl::read_excel(f, sheet = xls_names[1], na = c("", "-")) |>
    rename(MWh = `Hourly Demand Met (in MW)`) |>
    # filter out ending-info-rows
    filter(!grepl("NITI|https", MWh)) |>
    filter(MWh != "") |>
    # optional filter out "Totals"
    filter(!grepl("Total", MWh)) |>
    mutate(MWh = as.numeric(MWh)) |>
    # Split "State" and "year"
    mutate(
      # date = as.Date(paste0(year, "-", month, "-", "15")), # assuming mid-month
      # source = sub(" - Generation", "", source) |> trimws(),
      year =  sub(".* - (\\d+)$", "\\1", State) |> trimws(),
      datetime = paste(Date, year, sep = ", "),
      datetime = as.POSIXct(datetime, format = "%d-%b %I%p, %Y", tz = TZ),
      State = sub(" - \\d+$", "", State),
      .before = 1
    ) |>
    as.data.table()
}

# Merge the list into one table
load_curve_2025 <- rbindlist(ll)

load_curve_2025 <- load_curve_2025 |>
  # group_by(datetime, state) |>
  # summarise(GWh = sum(GWh))
  mutate(
    DATE = as.Date(datetime),
    HOUR = hour(datetime),
    YDAY = yday(datetime),
    MONTH = month(datetime) |> 
      factor(levels = 1:12, ordered = TRUE),
    MONTH_abb = month.abb[MONTH] |> 
      factor(levels = month.abb, ordered = TRUE),
    SEASON = case_when(
      MONTH %in% c(12, 1, 2) ~ "Winter",
      MONTH %in% (3:5) ~ "Spring",
      MONTH %in% (6:8) ~ "Summer",
      MONTH %in% (9:11) ~ "Autumn"
    ) |> 
      factor(ordered = TRUE, 
             levels = c("Spring", "Summer", "Autumn", "Winter")),
    QUARTER = case_when(
      MONTH %in% (4:6) ~ "Q1",
      MONTH %in% (6:8) ~ "Q2",
      MONTH %in% (9:12) ~ "Q3",
      MONTH %in% (1:3) ~ "Q4"
    ),
    WEEKDAY = format(datetime, "%a") |>
      factor(ordered = TRUE,
             levels = format(as.Date("1970-01-04") + 0:6, "%a"))
  ) |>
  as.data.table()  

save(load_curve_2025, file = "data/load_curve_2025.RData")

```

### Plot loadcurve

```{r}
p <- load_curve_2025 |>
  mutate(YM = paste(year, MONTH, sep = "-")) |>
  group_by(YM, State) |>
  summarise(GWh = sum(MWh/1e3, na.rm = T), .groups = "drop") |>
  ggplot() +
  geom_col(aes(YM, GWh/1e3, fill = GWh/1e4)) +
  scale_fill_viridis_c(name = "TWh", direction = 1, option = "H") +
  labs(x = "", y = "TWh") +
  facet_wrap(~State, scales = "free_y") +
  theme_bw()
# p
ggplotly(p)

```

```{r}
p <- load_curve_2025 |>
  filter(grepl("Mah", State), year == "2024") |>
  ggplot() +
  geom_line(aes(HOUR, MWh, colour = MONTH_abb, group = YDAY), 
            # linewidth = .5, 
            alpha = .75) +
  # scale_colour_brewer(palette = "Paired", name = "Month") +
  scale_colour_viridis_d(option = "H", alpha = .75, name = "Month") +
  # scale_colour_viridis_d(option = "H", alpha = .75, limits = c(0, 12),
                         # name = "Month") +
  facet_wrap(~State, scales = "free_y") +
  labs(x = "Hour of Day", y = "GWh") +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(linewidth = 1.5)))
# p
ggsave("images/load_curve_2024.png", p, scale = 1.5, width = 10, 
       height = 8)


```

```{r}
p <- load_curve_2025 |>
  # filter(year == 2024) |>
  group_by(year, datetime, DATE, HOUR, YDAY) |>
  summarise(GWh = sum(MWh/1e3, na.rm = T), .groups = "drop") |>
  as.data.table() |>
  ggplot() +
  geom_tile(aes(DATE, HOUR, fill = GWh)) +
  scale_fill_viridis_c(option = "H", name = "GWh") +
  scale_x_date(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 24, by = 4)) +  
  # facet_wrap(~scenario, scales = "free_y") +
  # facet_wrap(~region, scales = "free") +
  labs(x = "", y = "Hour of Day") +
  theme_bw()

ggsave("images/load_curve_heatmap_2024-25.png", p, scale = 1.5, width = 10, 
       height = 5)


```

## Map states to reg32 and reg5
```{r, eval=FALSE}
NITI_Aayog_states <- gen_capacity_2025 |> select(State) |> unique()

states_to_regs <- add_region_codes(
  NITI_Aayog_states,
  state_col = "State",
  target_level = 35,
  add_names = T,
  show_scores = TRUE  # See which aggregation level was used
)

fwrite(states_to_regs, "data/states_to_regs.csv")

```



