---
title: "setup"
author: '@WRI IDEEA team'
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

Note: 
This document has a list of tools/scripts for the project setup and data (pre-)processing normally to run once during the creation and setting up the project's environment/workspace. 


## Project setup
Installs IDEEA packages, checks if Julia is installed.

```{r}
source("R/ideea_installer.R")
ideea_install_check()
check_julia()

```

## Reset utility
A utility for a soft or hard reset of R environment. Run it (optionally) in every session start to ensure no unexpected data or packages are loaded and potentially conflict with key packages. 
```{r}
# usage:
source("R/reset.R")
reset_session() # with prompting a dialogue
# or
reset_session(prompt = FALSE)
```

## Copying saved scenarios from another computer
If a scenario has been solved and saved on another computer, the paths in the scenario object need to be updated.
It can be done with the function `update_scenario_paths()`. 
Currently we have two scenarios saved on the WRI sharepoint: `REF` and `CAP`. 
The following code assumes that the scenarios have been copied to `ideea_scenarios()` on the local computer.
The path adjustment has to be done for each scenario object separately and the updated object has to be saved again as `scen.RData` in the respective scenario folder.
```{r}
library(IDEEA)
ideea_scenarios()
source("R/update_scenario_paths.R") # script with the update_scenario_paths() function

# Load the scenario objects
# REF scenario
load_scenario(
  file.path(ideea_scenarios(), "REF_IDEEA_ELC_reg5_d365_h24_Y2020_2070_by_20"),
  overwrite = TRUE # overwrite in memory (RAM) if already loaded
 )
# CAP scenario
load_scenario(
  file.path(ideea_scenarios(), "CAP_IDEEA_ELC_reg5_d365_h24_Y2020_2070_by_20"),
  overwrite = TRUE
)

# Update paths in the scenario object
scen <- update_scenario_paths(
  .scen$REF,
  scenarios_old_dir = "scenarios/WRI",
  scenarios_new_dir = ideea_scenarios() # !!! adjust to your local path
)
scen@path # check if the path has been updated
# check if the data is accessible
getData(scen, "vObjective")
if (length(getData(scen, "vObjective")) > 0) {
  message("Data access check successful!")
  # Save the updated scenario object again as scen.RData in the respective scenario folder
  save(scen, file = file.path(scen@path, "scen.RData"))
} else {
  stop("Data access check failed!")
}

# Do the same for the CAP scenario
scen <- update_scenario_paths(
  .scen$CAP,
  scenarios_old_dir = "scenarios/WRI",
  scenarios_new_dir = ideea_scenarios()
)
# check if the data is accessible
getData(scen, "vObjective")
if (length(getData(scen, "vObjective")) > 0) {
  message("Data access check successful!")
  # Save the updated scenario object again as scen.RData in the respective scenario folder
  save(scen, file = file.path(scen@path, "scen.RData"))
} else {
  stop("Data access check failed!")
}

## Re-load the updated scenario to check if everything works
load_scenario(
  file.path(ideea_scenarios(), "REF_IDEEA_ELC_reg5_d365_h24_Y2020_2070_by_20"),
  overwrite = TRUE
)
load_scenario(
  file.path(ideea_scenarios(), "CAP_IDEEA_ELC_reg5_d365_h24_Y2020_2070_by_20"),
  overwrite = TRUE
)
sns <- list(.scen$REF, .scen$CAP)
getData(sns, "vObjective")

```

## Study list (minimal requirements)

A list of cheat sheets for broadly used R-packages:
https://posit.co/resources/cheatsheets/ 

Recommended (used in the project):  

* [Data transformation with dplyr](https://rstudio.github.io/cheatsheets/html/data-transformation.html)  
`dplyr` offers an intuitive way to work with tabular data. It's "macro-language" is adopted by multiple data formats and packages. We use it all the time.  

* [Data visualization with ggplot2](https://rstudio.github.io/cheatsheets/html/data-visualization.html)
Most of figures in the project are based on `ggplot2` package.

* [Publish and Share with Quarto](https://rstudio.github.io/cheatsheets/html/quarto.html)  
`quarto` is a successor of R-markdown to create reproducible research documents, presentations, web-pages.  


