---
title: "setup"
author: '@WRI IDEEA team'
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project setup

```{r}
source("R/ideea_installer.R")
ideea_install_check()
check_julia()

```


## Copying saved scenarios from another computer
If a scenario has been solved and saved on another computer, the paths in the scenario object need to be updated.
It can be done with the function `update_scenario_paths()`. 
Currently we have two scenarios saved on the WRI sharepoint: `REF` and `CAP`. 
The following code assumes that the scenarios have been copied to `ideea_scenarios()` on the local computer.
The path adjustment has to be done for each scenario object separately and the updated object has to be saved again as `scen.RData` in the respective scenario folder.
```{r, eval=FALSE}
library(IDEEA)
ideea_scenarios()
source("R/update_scenario_paths.R") # script with the update_scenario_paths() function

# Load the scenario objects
# REF scenario
load_scenario(
  file.path(ideea_scenarios(), "REF_IDEEA_ELC_reg5_d365_h24_Y2020_2070_by_20"),
  overwrite = TRUE # overwrite in memory (RAM) if already loaded
 )
# CAP scenario
load_scenario(
  file.path(ideea_scenarios(), "CAP_IDEEA_ELC_reg5_d365_h24_Y2020_2070_by_20"),
  overwrite = TRUE
)

# Update paths in the scenario object
scen <- update_scenario_paths(
  .scen$REF,
  scenarios_old_dir = "scenarios/WRI",
  scenarios_new_dir = ideea_scenarios() # !!! adjust to your local path
)
scen@path # check if the path has been updated
# check if the data is accessible
getData(scen, "vObjective")
if (length(getData(scen, "vObjective")) > 0) {
  message("Data access check successful!")
  # Save the updated scenario object again as scen.RData in the respective scenario folder
  save(scen, file = file.path(scen@path, "scen.RData"))
} else {
  stop("Data access check failed!")
}

# Do the same for the CAP scenario
scen <- update_scenario_paths(
  .scen$CAP,
  scenarios_old_dir = "scenarios/WRI",
  scenarios_new_dir = ideea_scenarios()
)
# check if the data is accessible
getData(scen, "vObjective")
if (length(getData(scen, "vObjective")) > 0) {
  message("Data access check successful!")
  # Save the updated scenario object again as scen.RData in the respective scenario folder
  save(scen, file = file.path(scen@path, "scen.RData"))
} else {
  stop("Data access check failed!")
}

## Re-load the updated scenario to check if everything works
load_scenario(
  file.path(ideea_scenarios(), "REF_IDEEA_ELC_reg5_d365_h24_Y2020_2070_by_20"),
  overwrite = TRUE
)
load_scenario(
  file.path(ideea_scenarios(), "CAP_IDEEA_ELC_reg5_d365_h24_Y2020_2070_by_20"),
  overwrite = TRUE
)
sns <- list(.scen$REF, .scen$CAP)
getData(sns, "vObjective")

```

