---
title: "IDEEA: transportation module"
editor_options: 
  chunk_output_type: console
output: html_document
params:
  # true → model refueling/charging infrastructure as INFR_* newTechnology objects
  # (with vehicle cinp2ainp linkage) instead of unconstrained newSupply (default).
  infra_as_tech: false
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.height = 5,
  fig.width = 7,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE}
# library(IDEEA)
# library(data.table)
source("R/setup.R")
source("R/ideea_report.r")
source("R/ideea_vehicle.R")
source("R/ideea_levcost.R")
source("R/read_transport_inputs.R")
# library(energyRt.dev)
```

Here we develop a transportation sector model for India. The structure
mirrors the **Energy Policy Simulator (EPS)** transport module, which
classifies vehicles by **type** (six categories) and **service**
(passenger / freight). Each vehicle type carries a set of fuel/powertrain
technologies (battery-electric, natural gas, gasoline, diesel, plug-in
hybrid, LPG, hydrogen).

Branches modelled (matches EPS `Number of Vehicles` taxonomy and the
`build_branch_repo()` calls below):

| EPS vehicle type | Passenger branch    | Freight branch      |
|------------------|---------------------|---------------------|
| LDV              | `LDV_psgr`          | `LDV_frgt`          |
| HDV              | `HDV_psgr`          | `HDV_frgt`          |
| Aircraft         | `aircraft_psgr`     | `aircraft_frgt`     |
| Rail             | `rail_psgr`         | `rail_frgt`         |
| Ships            | `ships_psgr`        | `ships_frgt`        |
| Motorbikes       | `motorbikes_psgr`   | `motorbikes_frgt`   |

EPS reference data (vehicle stocks, new-vehicle prices, fuel economies,
maintenance costs, etc.) lives in
`Files for tranport sector reference - IDEEA/` as one folder per EPS
variable code (e.g. `SYVbT/` = *Start Year Vehicles by Technology*,
`BNVP/` = *Baseline New Vehicle Price*). These feed the consolidated
input workbook `data/transport_inputs.xlsx`, which is read once by
`read_transport_inputs()` and then sliced per-branch by
`build_branch_repo()`.

All branches share a common set of fuel commodities (GSL, DSL, CNG, LPG,
ELCPJ, H2, plus blends) and emission commodities (CO2, PM).


## Road transport



## Light Duty Vehicles (LDV)  

LDV or cars are modeled to deliver two types of transportation services, based on efficiency/fuel consumption per trip:  
* highway passenger*km (`P-LDV-HWY`)   
* city passenger*km (`P-LDV-CTY`) 


### Inputs

```{r shared_setup}
tabs <- read_transport_inputs("data/transport_inputs.xlsx")
options(infra_as_tech = isTRUE(params$infra_as_tech))  # propagates to all build_branch_repo() calls
names(tabs)

set_default_solver(solver_options$glpk)
```

---

## LDV Passenger model

```{r ldv_psgr_build, results='hide'}
repo_LDV_psgr <- build_branch_repo(tabs, "LDV_psgr")
repo_LDV_psgr <- add(repo_LDV_psgr,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_LDV_psgr)
```

```{r ldv_psgr_model}
model_LDV_psgr <- newModel(
  name     = "LDV_psgr",
  desc     = "LDV passenger transport model",
  data     = repo_LDV_psgr,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r ldv_psgr_solve}
scen_LDV_psgr_BAU <- interpolate(
  model_LDV_psgr,
  name = "LDV_psgr_BAU",
  desc = "LDV passenger BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_LDV_psgr_BAU <- write_sc(scen_LDV_psgr_BAU)
scen_LDV_psgr_BAU <- solve_scenario(scen_LDV_psgr_BAU)
scen_LDV_psgr_BAU <- read_solution(scen_LDV_psgr_BAU)
summary(scen_LDV_psgr_BAU)
scen_LDV_psgr_BAU <- save_scenario(scen_LDV_psgr_BAU)

```

```{r ldv_psgr_results}
pDemand_LDV_psgr <- getData(
  scen_LDV_psgr_BAU,
  "pDemand", merge = T
  )

getData(scen_LDV_psgr_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
getData(scen_LDV_psgr_BAU, name = "vTechNewCap", merge = TRUE) 
  # pivot_wider(names_from = "tech")
```

```{r ldv_psgr_levcost, message=FALSE, warning=FALSE}
# ── Levelized cost comparison for all LDV-passenger technologies ─────────────
ldv_psgr_techs <- Filter(
  function(x) inherits(x, "technology"),
  repo_LDV_psgr@data
)

lc_LDV_psgr <- ideea_levcost(
  object   = ldv_psgr_techs,
  group    = "o",
  repo     = repo_LDV_psgr,
  discount = 0.05,
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5,
  calendar = newCalendar(),
  region   = "IND",
  verbose  = FALSE
)
```

```{r ldv_psgr_levcost_npv, fig.cap="Levelized cost comparison (NPV-weighted)"}
autoplot(lc_LDV_psgr, type = "npv")
```

```{r ldv_psgr_levcost_components, fig.cap="Levelized cost components by technology"}
autoplot(lc_LDV_psgr, type = "components")
```

```{r ldv_psgr_sanity, message=FALSE, eval=FALSE}
# ── Sanity check: 2024 fleet stock vs 2025 initial demand ────────────────────
# Unit chain: stock [1000 veh] × cap2act [Mveh-km / 1000-veh-yr]
#             × load [pax/veh] = MPkm/yr
fleet_t0 <- tabs[["fleet"]][
  tabs[["fleet"]][["branch"]] == "LDV_psgr" &
  tabs[["fleet"]][["year"]]   == 2024L, ]
tech_p <- tabs[["technologies"]][
  tabs[["technologies"]][["branch"]] == "LDV_psgr",
  c("name", "capacity_units", "avg_km_yr",
    "load_per_vehicle_hwy", "load_per_vehicle_cty")]

sanity <- merge(fleet_t0[, c("tech","stock")], tech_p,
                by.x = "tech", by.y = "name")
sanity$cap2act_Mvehkm  <- sanity$avg_km_yr * sanity$capacity_units / 1e6
hwy_share <- 0.20  # must match hwy_share_psgr in build_transport_inputs.R
sanity$service_MPkm <- round(
  sanity$stock * sanity$cap2act_Mvehkm *
  (hwy_share          * sanity$load_per_vehicle_hwy +
   (1 - hwy_share)    * sanity$load_per_vehicle_cty))

cat("=== LDV Passenger: fleet-to-service sanity check (2024 stock → MPkm) ===\n")
print(sanity[, c("tech","stock","cap2act_Mvehkm","service_MPkm")])
cat("\n  Total implied service from 2024 fleet:",
    sum(sanity$service_MPkm, na.rm = TRUE), "MPkm\n")

dem_t <- tabs[["demand"]][tabs[["demand"]][["branch"]] == "LDV_psgr", ]
yr_cols   <- grep("^y[0-9]{4}$", names(dem_t), value = TRUE)
dem_y2025 <- if ("y2025" %in% yr_cols) sum(dem_t$y2025, na.rm = TRUE) else NA_real_
cat("  Initial demand (2025):              ", round(dem_y2025), "MPkm\n")
cat("  Ratio (fleet service / demand):", round(sum(sanity$service_MPkm, na.rm=TRUE) /
                                               dem_y2025, 3), "\n")
cat("  (Ratio ~1 means stock is sized to meet demand; <<1 means demand >> existing fleet)\n")
```

---

## LDV Freight model

```{r ldv_frgt_build, results='hide'}
repo_LDV_frgt <- build_branch_repo(tabs, "LDV_frgt")
repo_LDV_frgt <- add(repo_LDV_frgt,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_LDV_frgt)
```

```{r ldv_frgt_model}
model_LDV_frgt <- newModel(
  name     = "LDV_frgt",
  desc     = "LDV freight transport model",
  data     = repo_LDV_frgt,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r ldv_frgt_solve}
scen_LDV_frgt_BAU <- interpolate(
  model_LDV_frgt,
  name = "LDV_frgt_BAU",
  desc = "LDV freight BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_LDV_frgt_BAU <- write_sc(scen_LDV_frgt_BAU)
scen_LDV_frgt_BAU <- solve_scenario(scen_LDV_frgt_BAU)
scen_LDV_frgt_BAU <- read_solution(scen_LDV_frgt_BAU)
summary(scen_LDV_frgt_BAU)
scen_LDV_frgt_BAU <- save_scenario(scen_LDV_frgt_BAU)
```

```{r ldv_frgt_results}
getData(scen_LDV_frgt_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

---

## HDV Passenger model

```{r hdv_psgr_build, results='hide'}
repo_HDV_psgr <- build_branch_repo(tabs, "HDV_psgr")
repo_HDV_psgr <- add(repo_HDV_psgr,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_HDV_psgr)
```

```{r hdv_psgr_model}
model_HDV_psgr <- newModel(
  name     = "HDV_psgr",
  desc     = "HDV passenger (bus) transport model",
  data     = repo_HDV_psgr,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r hdv_psgr_solve}
scen_HDV_psgr_BAU <- interpolate(
  model_HDV_psgr,
  name = "HDV_psgr_BAU",
  desc = "HDV passenger BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_HDV_psgr_BAU <- write_sc(scen_HDV_psgr_BAU)
scen_HDV_psgr_BAU <- solve_scenario(scen_HDV_psgr_BAU)
scen_HDV_psgr_BAU <- read_solution(scen_HDV_psgr_BAU)
summary(scen_HDV_psgr_BAU)
scen_HDV_psgr_BAU <- save_scenario(scen_HDV_psgr_BAU)
```

```{r hdv_psgr_results}
getData(scen_HDV_psgr_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

---

## HDV Freight model

```{r hdv_frgt_build, results='hide'}
repo_HDV_frgt <- build_branch_repo(tabs, "HDV_frgt")
repo_HDV_frgt <- add(repo_HDV_frgt,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_HDV_frgt)
```

```{r hdv_frgt_model}
model_HDV_frgt <- newModel(
  name     = "HDV_frgt",
  desc     = "HDV freight (truck) transport model",
  data     = repo_HDV_frgt,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r hdv_frgt_solve}
scen_HDV_frgt_BAU <- interpolate(
  model_HDV_frgt,
  name = "HDV_frgt_BAU",
  desc = "HDV freight BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_HDV_frgt_BAU <- write_sc(scen_HDV_frgt_BAU)
scen_HDV_frgt_BAU <- solve_scenario(scen_HDV_frgt_BAU)
scen_HDV_frgt_BAU <- read_solution(scen_HDV_frgt_BAU)
summary(scen_HDV_frgt_BAU)
scen_HDV_frgt_BAU <- save_scenario(scen_HDV_frgt_BAU)
```

```{r hdv_frgt_results}
getData(scen_HDV_frgt_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

---

## Aircraft Passenger model

```{r aircraft_psgr_build, results='hide'}
repo_aircraft_psgr <- build_branch_repo(tabs, "aircraft_psgr")
repo_aircraft_psgr <- add(repo_aircraft_psgr,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_aircraft_psgr)
```

```{r aircraft_psgr_model}
model_aircraft_psgr <- newModel(
  name     = "aircraft_psgr",
  desc     = "Aircraft passenger transport model",
  data     = repo_aircraft_psgr,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r aircraft_psgr_solve}
scen_aircraft_psgr_BAU <- interpolate(
  model_aircraft_psgr,
  name = "aircraft_psgr_BAU",
  desc = "Aircraft passenger BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_aircraft_psgr_BAU <- write_sc(scen_aircraft_psgr_BAU)
scen_aircraft_psgr_BAU <- solve_scenario(scen_aircraft_psgr_BAU)
scen_aircraft_psgr_BAU <- read_solution(scen_aircraft_psgr_BAU)
summary(scen_aircraft_psgr_BAU)
scen_aircraft_psgr_BAU <- save_scenario(scen_aircraft_psgr_BAU)
```

```{r aircraft_psgr_results}
getData(scen_aircraft_psgr_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

---

## Aircraft Freight model

```{r aircraft_frgt_build, results='hide'}
repo_aircraft_frgt <- build_branch_repo(tabs, "aircraft_frgt")
repo_aircraft_frgt <- add(repo_aircraft_frgt,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_aircraft_frgt)
```

```{r aircraft_frgt_model}
model_aircraft_frgt <- newModel(
  name     = "aircraft_frgt",
  desc     = "Aircraft freight transport model",
  data     = repo_aircraft_frgt,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r aircraft_frgt_solve}
scen_aircraft_frgt_BAU <- interpolate(
  model_aircraft_frgt,
  name = "aircraft_frgt_BAU",
  desc = "Aircraft freight BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_aircraft_frgt_BAU <- write_sc(scen_aircraft_frgt_BAU)
scen_aircraft_frgt_BAU <- solve_scenario(scen_aircraft_frgt_BAU)
scen_aircraft_frgt_BAU <- read_solution(scen_aircraft_frgt_BAU)
summary(scen_aircraft_frgt_BAU)
scen_aircraft_frgt_BAU <- save_scenario(scen_aircraft_frgt_BAU)
```

```{r aircraft_frgt_results}
getData(scen_aircraft_frgt_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

---

## Rail Passenger model

```{r rail_psgr_build, results='hide'}
repo_rail_psgr <- build_branch_repo(tabs, "rail_psgr")
repo_rail_psgr <- add(repo_rail_psgr,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_rail_psgr)
```

```{r rail_psgr_model}
model_rail_psgr <- newModel(
  name     = "rail_psgr",
  desc     = "Rail passenger transport model",
  data     = repo_rail_psgr,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r rail_psgr_solve}
scen_rail_psgr_BAU <- interpolate(
  model_rail_psgr,
  name = "rail_psgr_BAU",
  desc = "Rail passenger BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_rail_psgr_BAU <- write_sc(scen_rail_psgr_BAU)
scen_rail_psgr_BAU <- solve_scenario(scen_rail_psgr_BAU)
scen_rail_psgr_BAU <- read_solution(scen_rail_psgr_BAU)
summary(scen_rail_psgr_BAU)
scen_rail_psgr_BAU <- save_scenario(scen_rail_psgr_BAU)
```

```{r rail_psgr_results}
getData(scen_rail_psgr_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

---

## Rail Freight model

```{r rail_frgt_build, results='hide'}
repo_rail_frgt <- build_branch_repo(tabs, "rail_frgt")
repo_rail_frgt <- add(repo_rail_frgt,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_rail_frgt)
```

```{r rail_frgt_model}
model_rail_frgt <- newModel(
  name     = "rail_frgt",
  desc     = "Rail freight transport model",
  data     = repo_rail_frgt,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r rail_frgt_solve}
scen_rail_frgt_BAU <- interpolate(
  model_rail_frgt,
  name = "rail_frgt_BAU",
  desc = "Rail freight BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_rail_frgt_BAU <- write_sc(scen_rail_frgt_BAU)
scen_rail_frgt_BAU <- solve_scenario(scen_rail_frgt_BAU)
scen_rail_frgt_BAU <- read_solution(scen_rail_frgt_BAU)
summary(scen_rail_frgt_BAU)
scen_rail_frgt_BAU <- save_scenario(scen_rail_frgt_BAU)
```

```{r rail_frgt_results}
getData(scen_rail_frgt_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

---

## Ships Passenger model

```{r ships_psgr_build, results='hide'}
repo_ships_psgr <- build_branch_repo(tabs, "ships_psgr")
repo_ships_psgr <- add(repo_ships_psgr,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_ships_psgr)
```

```{r ships_psgr_model}
model_ships_psgr <- newModel(
  name     = "ships_psgr",
  desc     = "Ships passenger transport model",
  data     = repo_ships_psgr,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r ships_psgr_solve}
scen_ships_psgr_BAU <- interpolate(
  model_ships_psgr,
  name = "ships_psgr_BAU",
  desc = "Ships passenger BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_ships_psgr_BAU <- write_sc(scen_ships_psgr_BAU)
scen_ships_psgr_BAU <- solve_scenario(scen_ships_psgr_BAU)
scen_ships_psgr_BAU <- read_solution(scen_ships_psgr_BAU)
summary(scen_ships_psgr_BAU)
scen_ships_psgr_BAU <- save_scenario(scen_ships_psgr_BAU)
```

```{r ships_psgr_results}
getData(scen_ships_psgr_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

---

## Ships Freight model

```{r ships_frgt_build, results='hide'}
repo_ships_frgt <- build_branch_repo(tabs, "ships_frgt")
repo_ships_frgt <- add(repo_ships_frgt,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_ships_frgt)
```

```{r ships_frgt_model}
model_ships_frgt <- newModel(
  name     = "ships_frgt",
  desc     = "Ships freight transport model",
  data     = repo_ships_frgt,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r ships_frgt_solve}
scen_ships_frgt_BAU <- interpolate(
  model_ships_frgt,
  name = "ships_frgt_BAU",
  desc = "Ships freight BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_ships_frgt_BAU <- write_sc(scen_ships_frgt_BAU)
scen_ships_frgt_BAU <- solve_scenario(scen_ships_frgt_BAU)
scen_ships_frgt_BAU <- read_solution(scen_ships_frgt_BAU)
summary(scen_ships_frgt_BAU)
scen_ships_frgt_BAU <- save_scenario(scen_ships_frgt_BAU)
```

```{r ships_frgt_results}
getData(scen_ships_frgt_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

---

## Motorbikes Passenger model

```{r motorbikes_psgr_build, results='hide'}
repo_motorbikes_psgr <- build_branch_repo(tabs, "motorbikes_psgr")
repo_motorbikes_psgr <- add(repo_motorbikes_psgr,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_motorbikes_psgr)
```

```{r motorbikes_psgr_model}
model_motorbikes_psgr <- newModel(
  name     = "motorbikes_psgr",
  desc     = "Motorbikes passenger transport model",
  data     = repo_motorbikes_psgr,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r motorbikes_psgr_solve}
scen_motorbikes_psgr_BAU <- interpolate(
  model_motorbikes_psgr,
  name = "motorbikes_psgr_BAU",
  desc = "Motorbikes passenger BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_motorbikes_psgr_BAU <- write_sc(scen_motorbikes_psgr_BAU)
scen_motorbikes_psgr_BAU <- solve_scenario(scen_motorbikes_psgr_BAU)
scen_motorbikes_psgr_BAU <- read_solution(scen_motorbikes_psgr_BAU)
summary(scen_motorbikes_psgr_BAU)
scen_motorbikes_psgr_BAU <- save_scenario(scen_motorbikes_psgr_BAU)
```

```{r motorbikes_psgr_results}
getData(scen_motorbikes_psgr_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

---

## Motorbikes Freight model

```{r motorbikes_frgt_build, results='hide'}
repo_motorbikes_frgt <- build_branch_repo(tabs, "motorbikes_frgt")
repo_motorbikes_frgt <- add(repo_motorbikes_frgt,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_motorbikes_frgt)
```

```{r motorbikes_frgt_model}
model_motorbikes_frgt <- newModel(
  name     = "motorbikes_frgt",
  desc     = "Motorbikes freight transport model",
  data     = repo_motorbikes_frgt,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r motorbikes_frgt_solve}
scen_motorbikes_frgt_BAU <- interpolate(
  model_motorbikes_frgt,
  name = "motorbikes_frgt_BAU",
  desc = "Motorbikes freight BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_motorbikes_frgt_BAU <- write_sc(scen_motorbikes_frgt_BAU)
scen_motorbikes_frgt_BAU <- solve_scenario(scen_motorbikes_frgt_BAU)
scen_motorbikes_frgt_BAU <- read_solution(scen_motorbikes_frgt_BAU)
summary(scen_motorbikes_frgt_BAU)
scen_motorbikes_frgt_BAU <- save_scenario(scen_motorbikes_frgt_BAU)
```

```{r motorbikes_frgt_results}
getData(scen_motorbikes_frgt_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

---

## Combined transport model

All branches solved and validated individually — combine into one model.

```{r combined_build, results='hide'}
# Merge all 12 branch repos into one.
# Shared fuel commodities (GSL, DSL, …) are duplicated across branches —
# overwrite = TRUE keeps the last version without error.
repo_transport <- newRepository("repo_transport")
for (br_repo in list(
  repo_LDV_psgr, repo_LDV_frgt,
  repo_HDV_psgr, repo_HDV_frgt,
  repo_aircraft_psgr, repo_aircraft_frgt,
  repo_rail_psgr, repo_rail_frgt,
  repo_ships_psgr, repo_ships_frgt,
  repo_motorbikes_psgr, repo_motorbikes_frgt
)) {
  repo_transport <- add(repo_transport, br_repo@data, overwrite = TRUE)
}
names(repo_transport)
```

```{r combined_model}
model_transport <- newModel(
  name     = "transport",
  desc     = "Combined transport model — LDV + HDV passenger and freight",
  data     = repo_transport,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r combined_solve}
scen_transport_BAU <- interpolate(
  model_transport,
  name = "transport_BAU",
  desc = "Combined transport BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_transport_BAU <- write_sc(scen_transport_BAU)
scen_transport_BAU <- solve_scenario(scen_transport_BAU)
scen_transport_BAU <- read_solution(scen_transport_BAU)
summary(scen_transport_BAU)
scen_transport_BAU <- save_scenario(scen_transport_BAU)
```

```{r combined_results}
getData(scen_transport_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech") |> as.data.table()
getData(scen_transport_BAU, name = "vBalance", merge = TRUE, comm = "CO2")
```

## CO2 cap
```{r}
CO2_CAP_transport <- ideea_modules$electricity[[1]]$CO2_CAP
CO2_CAP_transport@rhs$rhs <- CO2_CAP_transport@rhs$rhs / 3
class(CO2_CAP_transport)

scen_transport_CAP <- interpolate(
  model_transport,
  CO2_CAP_transport,
  name = "transport_CAP",
  desc = "Combined transport CAP scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_transport_CAP <- write_sc(scen_transport_CAP)
scen_transport_CAP <- solve_scenario(scen_transport_CAP)
scen_transport_CAP <- read_solution(scen_transport_CAP)
summary(scen_transport_CAP)
scen_transport_CAP <- save_scenario(scen_transport_CAP)

```


```{r, eval=FALSE}
# Interactive explorer of scenarios
shiny::runApp("R/transport_explorer")
```













