---
title: "IDEEA: transportation module"
editor_options: 
  chunk_output_type: console
output: html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.height = 5,
  fig.width = 7,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE}
library(IDEEA)
library(data.table)
source("R/ideea_report.r")
# library(energyRt.dev)
```

Here we develop a transportation sector model for India with several sub-sectors or branches:  
* road transport  
  - light duty vehicles (LDV)  
  - medium duty vehicles (MDV)
  - heavy duty vehicles (HDV)  
  - buses   
    + city   
    + inter-city   
* rail transport  
* electric city transport  
  - trams and trolleys  
  - metro and short-distance trains  
* air transport  
  - domestic and international   
* water transport  
  - domestic  
  - international  

All transportation services deliver two types of services: passenger transportation, and freight. 


## Road transport



## Light Duty Vehicles (LDV)  

LDV or cars are modeled to deliver two types of transportation services, based on efficiency/fuel consumption per trip:  
* highway passenger*km (`P-LDV-HWY`)   
* city passenger*km (`P-LDV-CTY`) 


### Commodities
```{r}
GSL <- newCommodity(
  name = "GSL", 
  desc = "Gasoline (petrol)",
  unit = "PJ", 
  emis = data.frame(
    comm = "CO2",
    emis = 69.3 # tCO2/PJ, source: https://www.eia.gov/environment/emissions/co2_vol_mass.php
  ),
  timeframe = "ANNUAL")

DSL <- newCommodity(
  name = "DSL", 
  desc = "Diesel",
  unit = "PJ", 
  emis = data.frame(
    comm = "CO2",
    emis = 74.1 # tCO2/PJ, source: https://www.eia.gov/environment/emissions/co2_vol_mass.php
  ),
  timeframe = "ANNUAL")

BIO <- newCommodity(
  name = "BIO",
  desc = "Biofuels (ethanol, biodiesel)",
  unit = "PJ",
  timeframe = "ANNUAL",
  emis = data.frame(
    comm = "CO2",
    emis = 0 # tCO2/PJ, assuming carbon neutrality of biofuels
  )
)

ELCPJ <- newCommodity(
  name = "ELCPJ",
  unit = "PJ",
  timeframe = "ANNUAL"
)

# Blended fuel of 10% Ethanol and 90% gasoline
GSL10 <- newCommodity(
  name = "GSL10",
  desc = "Blended fuel of 10% Ethanol and 90% gasoline",
  unit = "PJ",
  timeframe = "ANNUAL",
  emis = data.frame(
    comm = "CO2",
    emis = 0.9 * 69.3 # tCO2/PJ, assuming 10% reduction in emissions compared to pure gasoline
  )
)

# Emissions
PM <- ideea_modules$electricity$reg5$repo_comm@data$PM
CO2 <- ideea_modules$electricity$reg5$repo_comm@data$CO2

```

### Transport services
```{r}
PLDVHWY <- newCommodity(
  name = "PLDVHWY", 
  desc = "LDV passengers transportation on highway",
  unit = "MPKm", # "million passenger * km", 
  timeframe = "ANNUAL")

PLDVCTY <- newCommodity(
  name = "PLDVCTY", 
  desc = "LDV passengers transportation in city",
  unit = "MPKm", # "million passenger * km", 
  timeframe = "ANNUAL")
```


### Demand
```{r}
DEM_PLDVHWY <- newDemand(
  name = "DEM_PLDVHWY",
  desc = "Demand for LDV passengers transportation on highway",
  commodity = "PLDVHWY",
  unit = "MPKm", # "million passenger * km", 
  region = "IND",
  dem = data.frame(
    region = "IND",
    year = 2024,
    dem = 1000 # million passenger * km a year
  )
)

DEM_PLDVCTY <- newDemand(
  name = "DEM_PLDVCTY",
  desc = "Demand for LDV passengers transportation in city",
  commodity = "PLDVCTY",
  unit = "MPKm", # "million passenger * km", 
  region = "IND",
  dem = data.frame(
    region = "IND",
    year = 2024,
    dem = 9000 # million passenger * km a year
  )
)

```

### Supply

```{r}
SUP_GSL <- newSupply(
  name = "SUP_GSL",
  desc = "Supply of gasoline (petrol)",
  commodity = "GSL",
  region = "IND",
  availability = data.frame(
    region = "IND",
    year = 2024,
    cost = 50 # MUSD/PJ
  )
)

SUP_DSL <- newSupply(
  name = "SUP_DSL",
  desc = "Supply of diesel",
  commodity = "DSL",
  region = "IND",
  availability = data.frame(
    region = "IND",
    year = 2024,
    cost = 40 # MUSD/PJ
  )
)

SUP_BIO <- newSupply(
  name = "SUP_BIO",
  desc = "Supply of biofuels (ethanol, biodiesel)",
  commodity = "BIO",
  region = "IND",
  availability = data.frame(
    region = "IND",
    year = 2024,
    cost = 70 # MUSD/PJ
  )
)

SUP_ELCPJ <- newSupply(
  name = "SUP_ELCPJ",
  desc = "Supply of electricity for transportation",
  commodity = "ELCPJ",
  region = "IND",
  availability = data.frame(
    region = "IND",
    year = 2024,
    cost = 30 # MUSD/PJ
  )
)

SUP_GSL10 <- newSupply(
  name = "SUP_GSL10",
  desc = "Supply of blended fuel of 10% Ethanol and 90% gasoline",
  commodity = "GSL10",
  region = "IND",
  availability = data.frame(
    region = "IND",
    year = 2024,
    cost = 60 # MUSD/PJ
  )
)


```


### Infrastructure

Q: Do we have a data on number of fuel stations, gasoline and diesel sales?

#### Services

Refuel infrastructure services by type, modeled as commodities.

```{r}
PETRO_REFUEL <- newCommodity(
  name = "PETRO_REFUEL", 
  desc = "Petrol refueling station services, gasoline and disel",
  unit = "PJ", # number of cars to fuel a year
  timeframe = "ANNUAL")

CNG_REFUEL <- newCommodity(
  name = "CNG_REFUEL", 
  desc = "Compressed Natural Gas (CNG) refueling station services",
  unit = "PJ", # number of cars to fuel a year
  timeframe = "ANNUAL")

LPG_REFUEL <- newCommodity(
  name = "LPG_REFUEL", 
  desc = "LPG station",
  unit = "PJ", # number of cars to fuel a year
  timeframe = "ANNUAL")

EV_CHARGING <- newCommodity(
  name = "EV_CHARGING", 
  desc = "Electric car charing station services",
  unit = "GWh", # number of cars to fuel a year
  timeframe = "ANNUAL")

```


#### Capacity

Refuel stations by type (Gas pumps, EV chargers, CNG and LPG stations) can be modeled as technology or supply objects. In case of technology, model will optimize required capacity of every service based on costs and technological parameters. In the simplest way, refueling services can be modeled as supply objects.
```{r}
# Supply case
SUP_PETRO_REFUEL <- newSupply(
  name = "SUP_PETRO_REFUEL",
  desc = "Petrol refueling station services, gasoline and disel",
  commodity = "PETRO_REFUEL",
  # unit = ... # either "PJ" or "1000 cars to fuel a year"
  availability = data.frame(
    # region = NA,
    year = 2024,
    # ava.up = 1000 # should correspond to the unit
    cost = 10 # MUSD/PJ
  )
  # region = NA_character_
)

SUP_CNG_REFUEL <- newSupply(
  name = "SUP_CNG_REFUEL",
  desc = "Compressed Natural Gas (CNG) refueling station services",
  commodity = "CNG_REFUEL",
  region = NA_character_,
  availability = data.frame(
    # region = NA,
    year = 2024,
    cost = 15 # MUSD/PJ
  )
)

SUP_LPG_REFUEL <- newSupply(
  name = "SUP_LPG_REFUEL",
  desc = "LPG station",
  commodity = "LPG_REFUEL",
  region = NA_character_,
  availability = data.frame(
    # region = NA,
    year = 2024,
    cost = 12 # MUSD/PJ
  )
)

SUP_EV_CHARGING <- newSupply(
  name = "SUP_EV_CHARGING",
  desc = "Electric car charing station services",
  commodity = "EV_CHARGING",
  region = NA_character_,
  availability = data.frame(
    # region = NA_character_,
    year = 2024,
    cost = 20 # MUSD/GWh
  )
)

energyRt.dev::draw(SUP_PETRO_REFUEL)

```



### Dual fuel cars
- Blended fuel 
- Dual fuel cars
```{r}


```



### Gasoline cars
```{r}
# Technologies
LDVG <- newTechnology(
  name = "LDVG",
  desc = "Gasoline Light Duty Vehicles",
  input = data.frame(
    comm = c("GSL", "BIO"),
    unit = "PJ", # for simplicity and electrification scenarios, normally "PJ"
    group = "i"
  ),
  output = data.frame(
    comm = c("PLDVHWY",  # passenger * km, highway
             "PLDVCTY"), # passenger * km, city
    unit = "MPKm", # million passenger * km
    group = "o" # a number or string, "o" = output
  ),
  units = list(
    capacity = "1000 Vehicles",
    activity = "million km, city", # measured in one of the output commodities
    costs = "MUSD"
  ),
  cap2act = 10, # 10000 km per car a year => 10000 * 1000 / 1e6
  ceff = data.frame(
    comm = c("GSL", "BIO", "PLDVHWY", "PLDVCTY"),
    use2cact = c(NA, NA, 3, 2), # !!! calculate million km per unit of input
    share.up = c(NA, .1, .4, .8),
    cact2cout = c(NA, NA, 2, 3) # passengers per vehicles (assumption)
  ),
  olife = list(olife = 10),
  capacity = list(
    region = NA,
    year = 2022
    # stock = 500
  ),
  invcost = list(
    invcost = 15 # 15000 [USD/car] * 1000 [cars] / 1e6 => MUSD/1000 cars
  ),
  fixom = list(
    fixom = .5 # repairs a year, taxes and fees, tires, oil change, parking etc.
  ),
  # varom = list(
  #   varom = ... # could be "tires, oil change" 
  # )
  aux = list(
    acomm = c("PETRO_REFUEL", "PM"), 
    unit = c("PJ", "kt")
  ), # N cars fueled a year
  aeff = list(
    acomm = c("PETRO_REFUEL", "PM"),
    cap2ainp = c(1, NA), # estimate number of trips to gas-station a year
    act2aout = c(NA, .01) # find emissions data per km
    )
  )

energyRt.dev::draw(LDVG)
draw(LDVG)
# ideea_report(LDVG, type = "vehicle", image_file = "images/LDV-gpt.png",
#              file = "tmp/LDVG_report_v5.pdf")

```


### Diesel cars
```{r}



```

### Hybrid cars
```{r}
LDV_HYB <- newTechnology(
  name = "LDV_HYB",
  desc = "Hybrid Light Duty Vehicles",
  input = data.frame(
    comm = c("GSL"),
    unit = "PJ", # for simplicity and electrification scenarios, normally "PJ"
    group = "i"
  ),
  output = data.frame(
    comm = c("PLDVHWY",  # passenger * km, highway
             "PLDVCTY"), # passenger * km, city
    unit = "MPKm", # million passenger * km
    group = "o" # a number or string, "o" = output
  ),
  units = list(
    capacity = "1000 Vehicles",
    activity = "million km, city", # measured in one of the output commodities
    costs = "MUSD"
  ),
  cap2act = 10, # 10000 km per car a year => 10000 * 1000 / 1e6
  ceff = data.frame(
    comm = c("GSL", "PLDVHWY", "PLDVCTY"),
    share.up = c(NA, .4, .8),
    cact2cout = c(NA, 2, 3) # passengers per
  ),
  olife = list(olife = 10),
  capacity = list(
    region = NA,
    year = 2022
    # stock = 500
  ),
  invcost = list(
    invcost = 20 # MUSD/1000 cars
  ),
  fixom = list(
    fixom = .5 # repairs a year, taxes and fees, tires, oil change, parking etc.
  ),
  aux = list(
    acomm = c("PETRO_REFUEL", "PM"), 
    unit = c("PJ", "kt")
  ), # N cars fueled a year
  aeff = list(
    acomm = c("PETRO_REFUEL", "PM"),
    cap2ainp = c(1, NA), # estimate number of trips to gas-station a year
    act2aout = c(NA, .01) # find emissions data per km
  )
)

energyRt.dev::draw(LDV_HYB)


```

### Plugged in hybrid cars
```{r}
LDV_PHEV <- newTechnology(
  name = "LDV_PHEV",
  desc = "Plugged in Hybrid Light Duty Vehicles",
  input = data.frame(
    comm = c("GSL", "ELCPJ"),
    unit = "PJ", # for simplicity and electrification scenarios, normally "PJ"
    group = "i"
  ),
  output = data.frame(
    comm = c("PLDVHWY",  # passenger * km, highway
             "PLDVCTY"), # passenger * km, city
    unit = "MPKm", # million passenger * km
    group = "o" # a number or string, "o" = output
  ),
  units = list(
    capacity = "1000 Vehicles",
    activity = "million km, city", # measured in one of the output commodities
    costs = "MUSD"
  ),
  cap2act = 10, # 10000 km per car a year =>
  ceff = data.frame(
    comm = c("GSL", "ELCPJ", "PLDVHWY", "PLDVCTY"),
    share.up = c(NA, NA, .4, .8),
    cact2cout = c(NA, NA, 2, 3) # passengers per
  ),
  olife = list(olife = 10),
  capacity = list(
    region = NA,
    year = 2022
    # stock = 100
  ),
  invcost = list(
    invcost = 25 # MUSD/1000 cars
  ),
  fixom = list(
    fixom = .5 # repairs a year, taxes and fees, tires, oil change
  ),
  aux = list(
    acomm = c("PETRO_REFUEL", "EV_CHARGING", "PM"),
    unit = c("PJ", "PJ", "kt")
  ), # N cars fueled a year
  aeff = list(
    acomm = c("PETRO_REFUEL", "EV_CHARGING", "PM"),
    comm = c(NA, "ELCPJ", NA),
    cap2ainp = c(1, NA, NA), # estimate number of trips to gas
    cinp2ainp = c(NA, 1, NA), # estimate number of trips to gas
    act2aout = c(NA, NA, .01) # find emissions data per km
  )
)

energyRt.dev::draw(LDV_PHEV)


```

### Electric cars
```{r}
LDV_EV <- newTechnology(
  name = "LDV_EV",
  desc = "Electric Light Duty Vehicles",
  input = data.frame(
    comm = c("ELCPJ"),
    unit = "PJ", # for simplicity and electrification scenarios, normally "PJ"
    group = "i"
  ),
  output = data.frame(
    comm = c("PLDVHWY",  # passenger * km, highway
             "PLDVCTY"), # passenger * km, city
    unit = "MPKm", # million passenger * km
    group = "o" # a number or string, "o" = output
  ),
  units = list(
    capacity = "1000 Vehicles",
    activity = "million km, city", # measured in one of the output commodities
    costs = "MUSD"
  ),
  cap2act = 10, # 10000 km per car a year =>
  ceff = data.frame(
    comm = c("ELCPJ", "PLDVHWY", "PLDVCTY"),
    share.up = c(NA, .4, .8),
    cact2cout = c(NA, 2, 3) # passengers per
  ),
  olife = list(olife = 10),
  capacity = list(
    region = NA,
    year = 2022
    # stock = 100
  ),
  invcost = list(
    invcost = 30 # MUSD/1000 cars
  ),
  fixom = list(
    fixom = .5 # repairs a year, taxes and fees, tires, oil change
  ),
  aux = list(
    acomm = c("EV_CHARGING", "PM"),
    unit = c("GWh", "kt")
  ), # N cars fueled a year
  aeff = list(
    acomm = c("EV_CHARGING", "PM"),
    cap2ainp = c(1, NA), # estimate number of trips to gas-station a year
    act2aout = c(NA, .01) # find emissions data per km
  )
)

energyRt.dev::draw(LDV_EV)

```

### Hydrogen cars
```{r}

```

### Model
```{r}
repo_LDV <- newRepository(
  name = "repo_LDV",
  desc = "Light Duty Vehicles (LDV) transportation services",
  # commodities - final demand
  PLDVHWY, PLDVCTY,
  # commodities - fuels
  GSL, DSL, BIO, ELCPJ, GSL10,
  # commodities - emissions
  PM, CO2,
  # commodities - refuel services
  PETRO_REFUEL, 
  CNG_REFUEL, LPG_REFUEL, EV_CHARGING,
  # Supplies - fuels
  SUP_GSL, SUP_DSL, SUP_BIO, SUP_ELCPJ, SUP_GSL10,
  # supply of refuel services
  SUP_PETRO_REFUEL, 
  SUP_CNG_REFUEL, SUP_LPG_REFUEL, SUP_EV_CHARGING,
  # technologies
  LDVG, 
  LDV_HYB, LDV_PHEV, LDV_EV,
  # demand
  DEM_PLDVHWY,
  DEM_PLDVCTY
)

ldv_model <- newModel(
  name = "ldv_model",
  desc = "Light Duty Vehicles (LDV) transportation services model",
  data = repo_LDV,
  region = "IND",
  debug = data.frame(
    # comm = c("GSL", "BIO", "PLDVHWY", "PLDVCTY"),
    # dummyImport = c(1e5, 1e5, 1e7, 1e7)
    # dummyExport = c(NA, 1e5)
    dummyImport = c(1e5)
    # dummyExport = c(1e5)
  ),
  discount = 0.05,
  # calendar = ideea_modules$calendars$calendar_d365_h24,
  calendar = newCalendar(),
  horizon = ideea_modules$horizons$horizon_2020_2060_by_5
  # regions = c("IND"),
  # time_steps = c("2024"),
  # time_slices = c("ANNUAL"),
  # solver = "glpk"
)

```

### Reference scenario (BAU)
```{r}
set_default_solver(solver_options$glpk)

scen_LDV_BAU <- interpolate(
  ldv_model, 
  name = "LDV_BAU", 
  desc = "Reference scenario (BAU) for Light Duty Vehicles (LDV)")

scen_LDV_BAU <- write_sc(scen_LDV_BAU)

scen_LDV_BAU <- solve_scenario(scen_LDV_BAU)
scen_LDV_BAU <- read_solution(scen_LDV_BAU)

```

```{r}
getData(scen_LDV_BAU, name = "vTechCap", merge = T) |> 
  pivot_wider(names_from = "tech")

getData(scen_LDV_BAU, name = "vTechNewCap", merge = T) |> 
  pivot_wider(names_from = "tech")

```

### CO2 constraint
```{r}
getData(scen_LDV_BAU, name = "vBalance", merge = T, comm = "CO2") 

CO2_CAP <- newConstraint(
  name = "CO2_CAP",
  desc = "Cap on CO2 emissions from transportation",
  eq = "<=",
  rhs = list(
    year = c(2030, 2050, 2070),
    rhs = c(200000, 1000, 1e-5)
  ),
  for.each = list(year = 2030:2070), # Cap total emission
  variable = list(
    variable = "vBalance",
    for.sum = list(
      comm = "CO2",
      slice = NULL,
      region = NULL
    )
  ),
  defVal = Inf,
  interpolation = "inter.forth"
)


```

### Carbon CAP scenario
```{r}
scen_LDV_CAP <- interpolate(
  ldv_model, 
  name = "LDV_CAP", 
  desc = "Scenario with CO2 emissions cap for Light Duty Vehicles (LDV)",
  CO2_CAP
)

scen_LDV_CAP <- write_sc(scen_LDV_CAP)
scen_LDV_CAP <- solve_scenario(scen_LDV_CAP)
scen_LDV_CAP <- read_solution(scen_LDV_CAP)  

summary(scen_LDV_CAP)

getData(scen_LDV_CAP, name = "vBalance", merge = T, comm = "CO2")

getData(scen_LDV_CAP, name = "vTechNewCap", merge = T) |> 
  pivot_wider(names_from = "tech")

```














