---
title: "IDEEA: transportation module"
editor_options: 
  chunk_output_type: console
output: html_document
params:
  # true → model refueling/charging infrastructure as INFR_* newTechnology objects
  # (with vehicle cinp2ainp linkage) instead of unconstrained newSupply (default).
  infra_as_tech: true
  # infra_as_tech: false  # true → model infra as unconstrained SUP_INFR_* supply objects while still
  # linking vehicle fuel inputs to infra service commodities via cinp2ainp aeff rows.
  # Mutually exclusive with infra_as_tech (infra_as_tech takes precedence if both true).
  # infra_as_supply: true
  infra_as_supply: false
  # true → run ideea_levcost() chunks for all branches (can be slow).
  eval_levcost: false
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.height = 5,
  fig.width = 7,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE}
# library(IDEEA)
# library(data.table)
source("R/setup.R")
source("R/ideea_report.r")
source("R/ideea_vehicle.R")
source("R/ideea_levcost.R")
source("R/read_transport_inputs.R")
eval_levcost <- isTRUE(params$eval_levcost)
# library(energyRt.dev)
```

Here we develop a transportation sector model for India. The structure
mirrors the **Energy Policy Simulator (EPS)** transport module, which
classifies vehicles by **type** (six categories) and **service**
(passenger / freight). Each vehicle type carries a set of fuel/powertrain
technologies (battery-electric, natural gas, gasoline, diesel, plug-in
hybrid, LPG, hydrogen).

Branches modelled (matches EPS `Number of Vehicles` taxonomy and the
`build_branch_repo()` calls below):

| EPS vehicle type | Passenger branch    | Freight branch      |
|------------------|---------------------|---------------------|
| LDV              | `LDV_psgr`          | `LDV_frgt`          |
| HDV              | `HDV_psgr`          | `HDV_frgt`          |
| Aircraft         | `aircraft_psgr`     | `aircraft_frgt`     |
| Rail             | `rail_psgr`         | `rail_frgt`         |
| Ships            | `ships_psgr`        | `ships_frgt`        |
| Motorbikes       | `motorbikes_psgr`   | `motorbikes_frgt`   |

EPS reference data (vehicle stocks, new-vehicle prices, fuel economies,
maintenance costs, etc.) lives in
`Files for tranport sector reference - IDEEA/` as one folder per EPS
variable code (e.g. `SYVbT/` = *Start Year Vehicles by Technology*,
`BNVP/` = *Baseline New Vehicle Price*). These feed the consolidated
input workbook `data/transport_inputs.xlsx`, which is read once by
`read_transport_inputs()` and then sliced per-branch by
`build_branch_repo()`.

All branches share a common set of fuel commodities (GSL, DSL, CNG, LPG,
ELCPJ, H2, plus blends) and emission commodities (CO2, PM).


## Road transport



## Light Duty Vehicles (LDV)  

LDV or cars are modeled to deliver two types of transportation services, based on efficiency/fuel consumption per trip:  
* highway passenger*km (`P-LDV-HWY`)   
* city passenger*km (`P-LDV-CTY`) 


### Inputs

```{r shared_setup}
tabs <- read_transport_inputs("data/transport_inputs.xlsx")
options(infra_as_tech   = isTRUE(params$infra_as_tech),    # propagates to all build_branch_repo() calls
        infra_as_supply = isTRUE(params$infra_as_supply))
names(tabs)

set_default_solver(solver_options$glpk)
```

---

## LDV Passenger model

```{r ldv_psgr_build, results='hide'}
repo_LDV_psgr <- build_branch_repo(tabs, "LDV_psgr")
repo_LDV_psgr <- add(repo_LDV_psgr,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_LDV_psgr)
repo_LDV_psgr@data$LDV_E10 |> draw()
repo_LDV_psgr@data$INFR_PETRO_REFUEL |> draw()
repo_LDV_psgr@data$SUP_CNG

# repo_LDV_psgr@data$INFR_PETRO_REFUEL |> design()
# repo_LDV_psgr@data$INFR_PETRO_REFUEL <- a

# repo_LDV_psgr@data$INFR_EV_CHARGING <- NULL
# repo_LDV_psgr@data$INFR_PETRO_REFUEL <- NULL
# repo_LDV_psgr@data$INFR_CNG_REFUEL <- NULL
# repo_LDV_psgr@data$INFR_LPG_REFUEL <- NULL

# repo_LDV_psgr@data$INFR_PETRO_REFUEL 

```

```{r ldv_psgr_model}
model_LDV_psgr <- newModel(
  name     = "LDV_psgr",
  desc     = "LDV passenger transport model",
  data     = repo_LDV_psgr,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r ldv_psgr_solve}
scen_LDV_psgr_BAU <- energyRt::interpolate(
  model_LDV_psgr,
  name = "LDV_psgr_BAU",
  desc = "LDV passenger BAU scenario"
  # debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_LDV_psgr_BAU <- write_sc(scen_LDV_psgr_BAU)
scen_LDV_psgr_BAU <- solve_scenario(scen_LDV_psgr_BAU)
scen_LDV_psgr_BAU <- read_solution(scen_LDV_psgr_BAU)
summary(scen_LDV_psgr_BAU)
scen_LDV_psgr_BAU <- save_scenario(scen_LDV_psgr_BAU)

```

```{r ldv_psgr_results}
pDemand_LDV_psgr <- getData(
  scen_LDV_psgr_BAU,
  "pDemand", merge = T
  )

getData(scen_LDV_psgr_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
getData(scen_LDV_psgr_BAU, name = "vTechNewCap", merge = TRUE) 
  # pivot_wider(names_from = "tech")
```

```{r ldv_psgr_levcost, message=FALSE, warning=FALSE, eval=eval_levcost}
# ── Levelized cost comparison for all LDV-passenger technologies ─────────────
if (isTRUE(eval_levcost)) {
  ldv_psgr_techs <- Filter(
    function(x) inherits(x, "technology") && !grepl("INFR_", x@name),
    repo_LDV_psgr@data
  )
  ldv_psgr_techs |> names()

  lc_LDV_psgr <- ideea_levcost(
    object   = ldv_psgr_techs,
    group    = "o",
    repo     = repo_LDV_psgr,
    discount = 0.05,
    horizon  = ideea_modules$horizons$horizon_2020_2060_by_5,
    calendar = newCalendar(),
    region   = "IND",
    verbose  = FALSE
  )
}
```

```{r ldv_psgr_levcost_npv, fig.cap="Levelized cost comparison (NPV-weighted)", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_LDV_psgr, type = "npv")
```

```{r ldv_psgr_levcost_components, fig.cap="Levelized cost components by technology", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_LDV_psgr, type = "components")
```

---

## LDV Freight model

```{r ldv_frgt_build, results='hide'}
repo_LDV_frgt <- build_branch_repo(tabs, "LDV_frgt")
repo_LDV_frgt <- add(repo_LDV_frgt,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_LDV_frgt)
```

```{r ldv_frgt_model}
model_LDV_frgt <- newModel(
  name     = "LDV_frgt",
  desc     = "LDV freight transport model",
  data     = repo_LDV_frgt,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r ldv_frgt_solve}
scen_LDV_frgt_BAU <- interpolate(
  model_LDV_frgt,
  name = "LDV_frgt_BAU",
  desc = "LDV freight BAU scenario"
  # debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_LDV_frgt_BAU <- write_sc(scen_LDV_frgt_BAU)
scen_LDV_frgt_BAU <- solve_scenario(scen_LDV_frgt_BAU)
scen_LDV_frgt_BAU <- read_solution(scen_LDV_frgt_BAU)
summary(scen_LDV_frgt_BAU)
scen_LDV_frgt_BAU <- save_scenario(scen_LDV_frgt_BAU)
```

```{r ldv_frgt_results}
getData(scen_LDV_frgt_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

```{r ldv_frgt_levcost, message=FALSE, warning=FALSE, eval=eval_levcost}
# ── Levelized cost comparison for all LDV-freight technologies ───────────────
if (isTRUE(eval_levcost)) {
  ldv_frgt_techs <- Filter(
    function(x) inherits(x, "technology") && !grepl("INFR_", x@name),
    repo_LDV_frgt@data
  )
  ldv_frgt_techs |> names()

  lc_LDV_frgt <- ideea_levcost(
    object   = ldv_frgt_techs,
    group    = "o",
    repo     = repo_LDV_frgt,
    discount = 0.05,
    horizon  = ideea_modules$horizons$horizon_2020_2060_by_5,
    calendar = newCalendar(),
    region   = "IND",
    verbose  = FALSE
  )
}
```

```{r ldv_frgt_levcost_npv, fig.cap="Levelized cost comparison (NPV-weighted)", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_LDV_frgt, type = "npv")
```

```{r ldv_frgt_levcost_components, fig.cap="Levelized cost components by technology", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_LDV_frgt, type = "components")
```

---

## HDV Passenger model

```{r hdv_psgr_build, results='hide'}
repo_HDV_psgr <- build_branch_repo(tabs, "HDV_psgr")
repo_HDV_psgr <- add(repo_HDV_psgr,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_HDV_psgr)
```

```{r hdv_psgr_model}
model_HDV_psgr <- newModel(
  name     = "HDV_psgr",
  desc     = "HDV passenger (bus) transport model",
  data     = repo_HDV_psgr,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r hdv_psgr_solve}
scen_HDV_psgr_BAU <- interpolate(
  model_HDV_psgr,
  name = "HDV_psgr_BAU",
  desc = "HDV passenger BAU scenario"
  # debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_HDV_psgr_BAU <- write_sc(scen_HDV_psgr_BAU)
scen_HDV_psgr_BAU <- solve_scenario(scen_HDV_psgr_BAU)
scen_HDV_psgr_BAU <- read_solution(scen_HDV_psgr_BAU)
summary(scen_HDV_psgr_BAU)
scen_HDV_psgr_BAU <- save_scenario(scen_HDV_psgr_BAU)
```

```{r hdv_psgr_results}
getData(scen_HDV_psgr_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

```{r hdv_psgr_levcost, message=FALSE, warning=FALSE, eval=eval_levcost}
# ── Levelized cost comparison for all HDV-passenger technologies ─────────────
if (isTRUE(eval_levcost)) {
  hdv_psgr_techs <- Filter(
    function(x) inherits(x, "technology") && !grepl("INFR_", x@name),
    repo_HDV_psgr@data
  )
  hdv_psgr_techs |> names()

  lc_HDV_psgr <- ideea_levcost(
    object   = hdv_psgr_techs,
    group    = "o",
    repo     = repo_HDV_psgr,
    discount = 0.05,
    horizon  = ideea_modules$horizons$horizon_2020_2060_by_5,
    calendar = newCalendar(),
    region   = "IND",
    verbose  = FALSE
  )
}
```

```{r hdv_psgr_levcost_npv, fig.cap="Levelized cost comparison (NPV-weighted)", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_HDV_psgr, type = "npv")
```

```{r hdv_psgr_levcost_components, fig.cap="Levelized cost components by technology", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_HDV_psgr, type = "components")
```

---

## HDV Freight model

```{r hdv_frgt_build, results='hide'}
repo_HDV_frgt <- build_branch_repo(tabs, "HDV_frgt")
repo_HDV_frgt <- add(repo_HDV_frgt,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_HDV_frgt)
```

```{r hdv_frgt_model}
model_HDV_frgt <- newModel(
  name     = "HDV_frgt",
  desc     = "HDV freight (truck) transport model",
  data     = repo_HDV_frgt,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r hdv_frgt_solve}
scen_HDV_frgt_BAU <- interpolate(
  model_HDV_frgt,
  name = "HDV_frgt_BAU",
  desc = "HDV freight BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_HDV_frgt_BAU <- write_sc(scen_HDV_frgt_BAU)
scen_HDV_frgt_BAU <- solve_scenario(scen_HDV_frgt_BAU)
scen_HDV_frgt_BAU <- read_solution(scen_HDV_frgt_BAU)
summary(scen_HDV_frgt_BAU)
scen_HDV_frgt_BAU <- save_scenario(scen_HDV_frgt_BAU)
```

```{r hdv_frgt_results}
getData(scen_HDV_frgt_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

```{r hdv_frgt_levcost, message=FALSE, warning=FALSE, eval=eval_levcost}
# ── Levelized cost comparison for all HDV-freight technologies ───────────────
if (isTRUE(eval_levcost)) {
  hdv_frgt_techs <- Filter(
    function(x) inherits(x, "technology") && !grepl("INFR_", x@name),
    repo_HDV_frgt@data
  )
  hdv_frgt_techs |> names()

  lc_HDV_frgt <- ideea_levcost(
    object   = hdv_frgt_techs,
    group    = "o",
    repo     = repo_HDV_frgt,
    discount = 0.05,
    horizon  = ideea_modules$horizons$horizon_2020_2060_by_5,
    calendar = newCalendar(),
    region   = "IND",
    verbose  = FALSE
  )
}
```

```{r hdv_frgt_levcost_npv, fig.cap="Levelized cost comparison (NPV-weighted)", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_HDV_frgt, type = "npv")
```

```{r hdv_frgt_levcost_components, fig.cap="Levelized cost components by technology", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_HDV_frgt, type = "components")
```

---

## Aircraft Passenger model

```{r aircraft_psgr_build, results='hide'}
repo_aircraft_psgr <- build_branch_repo(tabs, "aircraft_psgr")
repo_aircraft_psgr <- add(repo_aircraft_psgr,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_aircraft_psgr)
```

```{r aircraft_psgr_model}
model_aircraft_psgr <- newModel(
  name     = "aircraft_psgr",
  desc     = "Aircraft passenger transport model",
  data     = repo_aircraft_psgr,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r aircraft_psgr_solve}
scen_aircraft_psgr_BAU <- interpolate(
  model_aircraft_psgr,
  name = "aircraft_psgr_BAU",
  desc = "Aircraft passenger BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_aircraft_psgr_BAU <- write_sc(scen_aircraft_psgr_BAU)
scen_aircraft_psgr_BAU <- solve_scenario(scen_aircraft_psgr_BAU)
scen_aircraft_psgr_BAU <- read_solution(scen_aircraft_psgr_BAU)
summary(scen_aircraft_psgr_BAU)
scen_aircraft_psgr_BAU <- save_scenario(scen_aircraft_psgr_BAU)
```

```{r aircraft_psgr_results}
getData(scen_aircraft_psgr_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

```{r aircraft_psgr_levcost, message=FALSE, warning=FALSE, eval=eval_levcost}
# ── Levelized cost comparison for all aircraft-passenger technologies ─────────
if (isTRUE(eval_levcost)) {
  aircraft_psgr_techs <- Filter(
    function(x) inherits(x, "technology") && !grepl("INFR_", x@name),
    repo_aircraft_psgr@data
  )
  aircraft_psgr_techs |> names()

  lc_aircraft_psgr <- ideea_levcost(
    object   = aircraft_psgr_techs,
    group    = "o",
    repo     = repo_aircraft_psgr,
    discount = 0.05,
    horizon  = ideea_modules$horizons$horizon_2020_2060_by_5,
    calendar = newCalendar(),
    region   = "IND",
    verbose  = FALSE
  )
}
```

```{r aircraft_psgr_levcost_npv, fig.cap="Levelized cost comparison (NPV-weighted)", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_aircraft_psgr, type = "npv")
```

```{r aircraft_psgr_levcost_components, fig.cap="Levelized cost components by technology", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_aircraft_psgr, type = "components")
```

---

## Aircraft Freight model

```{r aircraft_frgt_build, results='hide'}
repo_aircraft_frgt <- build_branch_repo(tabs, "aircraft_frgt")
repo_aircraft_frgt <- add(repo_aircraft_frgt,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_aircraft_frgt)
```

```{r aircraft_frgt_model}
model_aircraft_frgt <- newModel(
  name     = "aircraft_frgt",
  desc     = "Aircraft freight transport model",
  data     = repo_aircraft_frgt,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r aircraft_frgt_solve}
scen_aircraft_frgt_BAU <- interpolate(
  model_aircraft_frgt,
  name = "aircraft_frgt_BAU",
  desc = "Aircraft freight BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_aircraft_frgt_BAU <- write_sc(scen_aircraft_frgt_BAU)
scen_aircraft_frgt_BAU <- solve_scenario(scen_aircraft_frgt_BAU)
scen_aircraft_frgt_BAU <- read_solution(scen_aircraft_frgt_BAU)
summary(scen_aircraft_frgt_BAU)
scen_aircraft_frgt_BAU <- save_scenario(scen_aircraft_frgt_BAU)
```

```{r aircraft_frgt_results}
getData(scen_aircraft_frgt_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

```{r aircraft_frgt_levcost, message=FALSE, warning=FALSE, eval=eval_levcost}
# ── Levelized cost comparison for all aircraft-freight technologies ───────────
if (isTRUE(eval_levcost)) {
  aircraft_frgt_techs <- Filter(
    function(x) inherits(x, "technology") && !grepl("INFR_", x@name),
    repo_aircraft_frgt@data
  )
  aircraft_frgt_techs |> names()

  lc_aircraft_frgt <- ideea_levcost(
    object   = aircraft_frgt_techs,
    group    = "o",
    repo     = repo_aircraft_frgt,
    discount = 0.05,
    horizon  = ideea_modules$horizons$horizon_2020_2060_by_5,
    calendar = newCalendar(),
    region   = "IND",
    verbose  = FALSE
  )
}
```

```{r aircraft_frgt_levcost_npv, fig.cap="Levelized cost comparison (NPV-weighted)", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_aircraft_frgt, type = "npv")
```

```{r aircraft_frgt_levcost_components, fig.cap="Levelized cost components by technology", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_aircraft_frgt, type = "components")
```

---

## Rail Passenger model

```{r rail_psgr_build, results='hide'}
repo_rail_psgr <- build_branch_repo(tabs, "rail_psgr")
repo_rail_psgr <- add(repo_rail_psgr,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_rail_psgr)
```

```{r rail_psgr_model}
model_rail_psgr <- newModel(
  name     = "rail_psgr",
  desc     = "Rail passenger transport model",
  data     = repo_rail_psgr,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r rail_psgr_solve}
scen_rail_psgr_BAU <- interpolate(
  model_rail_psgr,
  name = "rail_psgr_BAU",
  desc = "Rail passenger BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_rail_psgr_BAU <- write_sc(scen_rail_psgr_BAU)
scen_rail_psgr_BAU <- solve_scenario(scen_rail_psgr_BAU)
scen_rail_psgr_BAU <- read_solution(scen_rail_psgr_BAU)
summary(scen_rail_psgr_BAU)
scen_rail_psgr_BAU <- save_scenario(scen_rail_psgr_BAU)
```

```{r rail_psgr_results}
getData(scen_rail_psgr_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

```{r rail_psgr_levcost, message=FALSE, warning=FALSE, eval=eval_levcost}
# ── Levelized cost comparison for all rail-passenger technologies ─────────────
if (isTRUE(eval_levcost)) {
  rail_psgr_techs <- Filter(
    function(x) inherits(x, "technology") && !grepl("INFR_", x@name),
    repo_rail_psgr@data
  )
  rail_psgr_techs |> names()

  lc_rail_psgr <- ideea_levcost(
    object   = rail_psgr_techs,
    group    = "o",
    repo     = repo_rail_psgr,
    discount = 0.05,
    horizon  = ideea_modules$horizons$horizon_2020_2060_by_5,
    calendar = newCalendar(),
    region   = "IND",
    verbose  = FALSE
  )
}
```

```{r rail_psgr_levcost_npv, fig.cap="Levelized cost comparison (NPV-weighted)", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_rail_psgr, type = "npv")
```

```{r rail_psgr_levcost_components, fig.cap="Levelized cost components by technology", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_rail_psgr, type = "components")
```

---

## Rail Freight model

```{r rail_frgt_build, results='hide'}
repo_rail_frgt <- build_branch_repo(tabs, "rail_frgt")
repo_rail_frgt <- add(repo_rail_frgt,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_rail_frgt)
```

```{r rail_frgt_model}
model_rail_frgt <- newModel(
  name     = "rail_frgt",
  desc     = "Rail freight transport model",
  data     = repo_rail_frgt,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r rail_frgt_solve}
scen_rail_frgt_BAU <- interpolate(
  model_rail_frgt,
  name = "rail_frgt_BAU",
  desc = "Rail freight BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_rail_frgt_BAU <- write_sc(scen_rail_frgt_BAU)
scen_rail_frgt_BAU <- solve_scenario(scen_rail_frgt_BAU)
scen_rail_frgt_BAU <- read_solution(scen_rail_frgt_BAU)
summary(scen_rail_frgt_BAU)
scen_rail_frgt_BAU <- save_scenario(scen_rail_frgt_BAU)
```

```{r rail_frgt_results}
getData(scen_rail_frgt_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

```{r rail_frgt_levcost, message=FALSE, warning=FALSE, eval=eval_levcost}
# ── Levelized cost comparison for all rail-freight technologies ───────────────
if (isTRUE(eval_levcost)) {
  rail_frgt_techs <- Filter(
    function(x) inherits(x, "technology") && !grepl("INFR_", x@name),
    repo_rail_frgt@data
  )
  rail_frgt_techs |> names()

  lc_rail_frgt <- ideea_levcost(
    object   = rail_frgt_techs,
    group    = "o",
    repo     = repo_rail_frgt,
    discount = 0.05,
    horizon  = ideea_modules$horizons$horizon_2020_2060_by_5,
    calendar = newCalendar(),
    region   = "IND",
    verbose  = FALSE
  )
}
```

```{r rail_frgt_levcost_npv, fig.cap="Levelized cost comparison (NPV-weighted)", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_rail_frgt, type = "npv")
```

```{r rail_frgt_levcost_components, fig.cap="Levelized cost components by technology", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_rail_frgt, type = "components")
```

---

## Ships Passenger model

```{r ships_psgr_build, results='hide'}
repo_ships_psgr <- build_branch_repo(tabs, "ships_psgr")
repo_ships_psgr <- add(repo_ships_psgr,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_ships_psgr)
```

```{r ships_psgr_model}
model_ships_psgr <- newModel(
  name     = "ships_psgr",
  desc     = "Ships passenger transport model",
  data     = repo_ships_psgr,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r ships_psgr_solve}
scen_ships_psgr_BAU <- interpolate(
  model_ships_psgr,
  name = "ships_psgr_BAU",
  desc = "Ships passenger BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_ships_psgr_BAU <- write_sc(scen_ships_psgr_BAU)
scen_ships_psgr_BAU <- solve_scenario(scen_ships_psgr_BAU)
scen_ships_psgr_BAU <- read_solution(scen_ships_psgr_BAU)
summary(scen_ships_psgr_BAU)
scen_ships_psgr_BAU <- save_scenario(scen_ships_psgr_BAU)
```

```{r ships_psgr_results}
getData(scen_ships_psgr_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

```{r ships_psgr_levcost, message=FALSE, warning=FALSE, eval=eval_levcost}
# ── Levelized cost comparison for all ships-passenger technologies ────────────
if (isTRUE(eval_levcost)) {
  ships_psgr_techs <- Filter(
    function(x) inherits(x, "technology") && !grepl("INFR_", x@name),
    repo_ships_psgr@data
  )
  ships_psgr_techs |> names()

  lc_ships_psgr <- ideea_levcost(
    object   = ships_psgr_techs,
    group    = "o",
    repo     = repo_ships_psgr,
    discount = 0.05,
    horizon  = ideea_modules$horizons$horizon_2020_2060_by_5,
    calendar = newCalendar(),
    region   = "IND",
    verbose  = FALSE
  )
}
```

```{r ships_psgr_levcost_npv, fig.cap="Levelized cost comparison (NPV-weighted)", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_ships_psgr, type = "npv")
```

```{r ships_psgr_levcost_components, fig.cap="Levelized cost components by technology", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_ships_psgr, type = "components")
```

---

## Ships Freight model

```{r ships_frgt_build, results='hide'}
repo_ships_frgt <- build_branch_repo(tabs, "ships_frgt")
repo_ships_frgt <- add(repo_ships_frgt,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_ships_frgt)
```

```{r ships_frgt_model}
model_ships_frgt <- newModel(
  name     = "ships_frgt",
  desc     = "Ships freight transport model",
  data     = repo_ships_frgt,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r ships_frgt_solve}
scen_ships_frgt_BAU <- interpolate(
  model_ships_frgt,
  name = "ships_frgt_BAU",
  desc = "Ships freight BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_ships_frgt_BAU <- write_sc(scen_ships_frgt_BAU)
scen_ships_frgt_BAU <- solve_scenario(scen_ships_frgt_BAU)
scen_ships_frgt_BAU <- read_solution(scen_ships_frgt_BAU)
summary(scen_ships_frgt_BAU)
scen_ships_frgt_BAU <- save_scenario(scen_ships_frgt_BAU)
```

```{r ships_frgt_results}
getData(scen_ships_frgt_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

```{r ships_frgt_levcost, message=FALSE, warning=FALSE, eval=eval_levcost}
# ── Levelized cost comparison for all ships-freight technologies ──────────────
if (isTRUE(eval_levcost)) {
  ships_frgt_techs <- Filter(
    function(x) inherits(x, "technology") && !grepl("INFR_", x@name),
    repo_ships_frgt@data
  )
  ships_frgt_techs |> names()

  lc_ships_frgt <- ideea_levcost(
    object   = ships_frgt_techs,
    group    = "o",
    repo     = repo_ships_frgt,
    discount = 0.05,
    horizon  = ideea_modules$horizons$horizon_2020_2060_by_5,
    calendar = newCalendar(),
    region   = "IND",
    verbose  = FALSE
  )
}
```

```{r ships_frgt_levcost_npv, fig.cap="Levelized cost comparison (NPV-weighted)", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_ships_frgt, type = "npv")
```

```{r ships_frgt_levcost_components, fig.cap="Levelized cost components by technology", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_ships_frgt, type = "components")
```

---

## Motorbikes Passenger model

```{r motorbikes_psgr_build, results='hide'}
repo_motorbikes_psgr <- build_branch_repo(tabs, "motorbikes_psgr")
repo_motorbikes_psgr <- add(repo_motorbikes_psgr,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_motorbikes_psgr)
```

```{r motorbikes_psgr_model}
model_motorbikes_psgr <- newModel(
  name     = "motorbikes_psgr",
  desc     = "Motorbikes passenger transport model",
  data     = repo_motorbikes_psgr,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r motorbikes_psgr_solve}
scen_motorbikes_psgr_BAU <- interpolate(
  model_motorbikes_psgr,
  name = "motorbikes_psgr_BAU",
  desc = "Motorbikes passenger BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_motorbikes_psgr_BAU <- write_sc(scen_motorbikes_psgr_BAU)
scen_motorbikes_psgr_BAU <- solve_scenario(scen_motorbikes_psgr_BAU)
scen_motorbikes_psgr_BAU <- read_solution(scen_motorbikes_psgr_BAU)
summary(scen_motorbikes_psgr_BAU)
scen_motorbikes_psgr_BAU <- save_scenario(scen_motorbikes_psgr_BAU)
```

```{r motorbikes_psgr_results}
getData(scen_motorbikes_psgr_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

```{r motorbikes_psgr_levcost, message=FALSE, warning=FALSE, eval=eval_levcost}
# ── Levelized cost comparison for all motorbikes-passenger technologies ───────
if (isTRUE(eval_levcost)) {
  motorbikes_psgr_techs <- Filter(
    function(x) inherits(x, "technology") && !grepl("INFR_", x@name),
    repo_motorbikes_psgr@data
  )
  motorbikes_psgr_techs |> names()

  lc_motorbikes_psgr <- ideea_levcost(
    object   = motorbikes_psgr_techs,
    group    = "o",
    repo     = repo_motorbikes_psgr,
    discount = 0.05,
    horizon  = ideea_modules$horizons$horizon_2020_2060_by_5,
    calendar = newCalendar(),
    region   = "IND",
    verbose  = FALSE
  )
}
```

```{r motorbikes_psgr_levcost_npv, fig.cap="Levelized cost comparison (NPV-weighted)", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_motorbikes_psgr, type = "npv")
```

```{r motorbikes_psgr_levcost_components, fig.cap="Levelized cost components by technology", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_motorbikes_psgr, type = "components")
```

---

## Motorbikes Freight model

```{r motorbikes_frgt_build, results='hide'}
repo_motorbikes_frgt <- build_branch_repo(tabs, "motorbikes_frgt")
repo_motorbikes_frgt <- add(repo_motorbikes_frgt,
  ideea_modules$electricity$reg5$repo_comm@data$CO2,
  ideea_modules$electricity$reg5$repo_comm@data$PM
)
names(repo_motorbikes_frgt)
```

```{r motorbikes_frgt_model}
model_motorbikes_frgt <- newModel(
  name     = "motorbikes_frgt",
  desc     = "Motorbikes freight transport model",
  data     = repo_motorbikes_frgt,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```

```{r motorbikes_frgt_solve}
scen_motorbikes_frgt_BAU <- interpolate(
  model_motorbikes_frgt,
  name = "motorbikes_frgt_BAU",
  desc = "Motorbikes freight BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_motorbikes_frgt_BAU <- write_sc(scen_motorbikes_frgt_BAU)
scen_motorbikes_frgt_BAU <- solve_scenario(scen_motorbikes_frgt_BAU)
scen_motorbikes_frgt_BAU <- read_solution(scen_motorbikes_frgt_BAU)
summary(scen_motorbikes_frgt_BAU)
scen_motorbikes_frgt_BAU <- save_scenario(scen_motorbikes_frgt_BAU)
```

```{r motorbikes_frgt_results}
getData(scen_motorbikes_frgt_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech")
```

```{r motorbikes_frgt_levcost, message=FALSE, warning=FALSE, eval=eval_levcost}
# ── Levelized cost comparison for all motorbikes-freight technologies ─────────
if (isTRUE(eval_levcost)) {
  motorbikes_frgt_techs <- Filter(
    function(x) inherits(x, "technology") && !grepl("INFR_", x@name),
    repo_motorbikes_frgt@data
  )
  motorbikes_frgt_techs |> names()

  lc_motorbikes_frgt <- ideea_levcost(
    object   = motorbikes_frgt_techs,
    group    = "o",
    repo     = repo_motorbikes_frgt,
    discount = 0.05,
    horizon  = ideea_modules$horizons$horizon_2020_2060_by_5,
    calendar = newCalendar(),
    region   = "IND",
    verbose  = FALSE
  )
}
```

```{r motorbikes_frgt_levcost_npv, fig.cap="Levelized cost comparison (NPV-weighted)", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_motorbikes_frgt, type = "npv")
```

```{r motorbikes_frgt_levcost_components, fig.cap="Levelized cost components by technology", eval=eval_levcost}
if (isTRUE(eval_levcost)) autoplot(lc_motorbikes_frgt, type = "components")
```

---

## Combined transport model

All branches solved and validated individually — combine into one model.

```{r combined_build, results='hide'}
# Merge all 12 branch repos into one.
# Shared fuel commodities (GSL, DSL, …) are duplicated across branches —
# overwrite = TRUE keeps the last version without error.
repo_transport <- newRepository("repo_transport")
for (br_repo in list(
  repo_LDV_psgr, repo_LDV_frgt,
  repo_HDV_psgr, repo_HDV_frgt,
  repo_aircraft_psgr, repo_aircraft_frgt,
  repo_rail_psgr, repo_rail_frgt,
  repo_ships_psgr, repo_ships_frgt,
  repo_motorbikes_psgr, repo_motorbikes_frgt
)) {
  repo_transport <- add(repo_transport, br_repo@data, overwrite = TRUE)
}
names(repo_transport)
```

```{r combined_model}
model_transport <- newModel(
  name     = "transport",
  desc     = "Combined transport model — LDV + HDV passenger and freight",
  data     = repo_transport,
  region   = "IND",
  debug    = data.frame(dummyImport = 1e5),
  discount = 0.05,
  calendar = newCalendar(),
  horizon  = ideea_modules$horizons$horizon_2020_2060_by_5
)
```


## Calibration run
```{r}
scen_transport_BAU_BY <- interpolate(
  model_transport,
  name = "transport_BAU_BY",
  desc = "Combined transport BAU scenario, base year calibration",
  horizon = newHorizon(2020)
)
scen_transport_BAU_BY <- write_sc(scen_transport_BAU_BY)
scen_transport_BAU_BY <- solve_scenario(scen_transport_BAU_BY)
scen_transport_BAU_BY <- read_solution(scen_transport_BAU_BY)
summary(scen_transport_BAU_BY)
scen_transport_BAU_BY <- save_scenario(scen_transport_BAU_BY)


```



```{r combined_solve}
scen_transport_BAU <- interpolate(
  model_transport,
  name = "transport_BAU",
  desc = "Combined transport BAU scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_transport_BAU <- write_sc(scen_transport_BAU)
scen_transport_BAU <- solve_scenario(scen_transport_BAU)
scen_transport_BAU <- read_solution(scen_transport_BAU)
summary(scen_transport_BAU)
scen_transport_BAU <- save_scenario(scen_transport_BAU)
```

```{r combined_results}
getData(scen_transport_BAU, name = "vTechCap", merge = TRUE) |>
  pivot_wider(names_from = "tech") |> as.data.table()
getData(scen_transport_BAU, name = "vBalance", merge = TRUE, comm = "CO2")
```

## CO2 cap
```{r}
CO2_CAP_transport <- ideea_modules$electricity[[1]]$CO2_CAP
CO2_CAP_transport@rhs$rhs <- CO2_CAP_transport@rhs$rhs / 3
class(CO2_CAP_transport)

scen_transport_CAP <- interpolate(
  model_transport,
  CO2_CAP_transport,
  name = "transport_CAP",
  desc = "Combined transport CAP scenario",
  debug = data.frame(dummyImport = 1e5, dummyExport = 1e5)
)
scen_transport_CAP <- write_sc(scen_transport_CAP)
scen_transport_CAP <- solve_scenario(scen_transport_CAP)
scen_transport_CAP <- read_solution(scen_transport_CAP)
summary(scen_transport_CAP)
scen_transport_CAP <- save_scenario(scen_transport_CAP)

```

## Explorer
```{r, eval=FALSE}
# Interactive explorer of scenarios
shiny::runApp("R/transport_explorer")
```













